---
title: "SusieR Cis and Trans Susie Fine-Mapping"
author: "Claudia Krueger"
date: "2025-07-01"
output: html_document
---

**Cis SusieR (~/ancestry_specific_folder/cis_susie.R)**
```{r}
library(susieR)
library(data.table)
library(argparse)

#setup command line argument parsing
parser <- ArgumentParser()
parser$add_argument("--oid", required=TRUE, help="OID number")
parser$add_argument("--chr", required=TRUE, help="chromosome number")
args <- parser$parse_args()

#define file paths
genotype_prefix <- "~/ancestry_specific_folder/ancestry_final"
phenotype_file <- "~/ancestry_specific_folder/tensor_lm_table.bed"
positions_file <- paste0("~/ancestry_specific_folder/0.05cis_locations.tsv")
output_prefix <- paste0("~/ancestry_specific_folder/cis_susieR/", args$oid, "_chr", args$chr)

#read in the position file
positions <- tryCatch({
  fread(positions_file, header=TRUE)
}, error = function(e) {
  stop(paste("Error reading positions file:", e$message))
})

#extract region of interest based on the specific OID and chromosome
current_position <- positions[positions$OlinkID == args$oid & positions$chr == paste0("chr", args$chr), ]

if(nrow(current_position) == 0) {
  stop(paste(args$oid, "on chromosome", args$chr, "not found in positions file"))
}

#get the min and max positions for this specific OID
min_pos <- max(0, current_position$left - 1000000)
max_pos <- current_position$right + 1000000
target_positions <- 0

#run PLINK2 command to make raw file of this region
geno_output <- paste0(output_prefix, "_genotypes")
cmd <- sprintf("plink2 --pfile %s --chr %s --from-bp %d --to-bp %d --export A --out %s", 
               genotype_prefix, args$chr, min_pos, max_pos, geno_output)
cat("Executing PLINK2 command:", cmd, "\n")
system_result <- system(cmd)

if (system_result != 0) {
  stop("PLINK2 extraction failed. Check PLINK log file for details.")
}

#check if file exists and read in extracted genotypes
geno_file <- paste0(geno_output, ".raw")
if (!file.exists(geno_file)) {
  stop("PLINK2 extraction failed. Output file not found.")
}

#read genotypes with error handling
genotypes <- tryCatch({
  fread(geno_file)
}, error = function(e) {
  stop(paste("Error reading genotype file:", e$message))
})

#remove metadata
meta_cols <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")
meta_cols <- meta_cols[meta_cols %in% colnames(genotypes)]
geno_cols <- setdiff(colnames(genotypes), meta_cols)

if (length(geno_cols) == 0) {
  stop("No genotype columns found in the PLINK output file.")
}

#find chr and pos of each SNP IDs
extract_pos <- function(snp_id) {
  if (is.na(snp_id)) return(NA)
  
  #if the snp_id column contains colons
  if (grepl(":", snp_id)) { 
    parts <- strsplit(snp_id, ":")[[1]]
    pos_part <- parts[2]
    #if the snp_id contains underscores and colons
    if (grepl("_", pos_part)) {
      pos <- as.numeric(strsplit(pos_part, "_")[[1]][1])
    } else {
      pos <- as.numeric(pos_part)
    }
    return(pos)
    #if the snp_id column contains underscores only
  } else if (grepl("_", snp_id)) {
    parts <- strsplit(snp_id, "_")[[1]]
    if (length(parts) >= 2) {
      return(as.numeric(parts[2]))
    }
  }
  return(NA) #if the position cannot be found
}

snp_positions <- sapply(geno_cols, extract_pos) #identify the positions
matching_snps <- geno_cols #use full genomic region

#create genotype matrix
X <- matrix(as.numeric(as.matrix(genotypes[, matching_snps, with=FALSE])), 
            nrow=nrow(genotypes), ncol=length(matching_snps))
colnames(X) <- matching_snps

#impute missing genotype data (shouldnt be any)
na_count <- sum(is.na(X))
if (na_count > 0) {
  cat("Imputing", na_count, "missing genotype values...\n")
  for (j in 1:ncol(X)) {
    na_idx <- is.na(X[,j])
    if (any(na_idx)) {
      X[na_idx, j] <- mean(X[!na_idx, j])
    }
  }
}

#read in protein level data
cat("Reading phenotype data...\n")
if (!file.exists(phenotype_file)) {
  stop(paste("Phenotype file not found:", phenotype_file))
}

#read phenotypes
pheno <- fread(phenotype_file, header=TRUE)

oid_rows <- which(pheno$protein_id == args$oid)
if (length(oid_rows) == 0) {
  stop(paste("protein", args$oid, "not found in phenotype file"))
}
oid_row <- oid_rows[1]

#remove any metadata
meta_cols <- c("#chr", "chr", "start", "end", "protein_id")
meta_cols <- meta_cols[meta_cols %in% colnames(pheno)]
sample_cols <- setdiff(colnames(pheno), meta_cols)

if (length(sample_cols) == 0) {
  stop("No sample columns found in phenotype file.")
}

#find protein level data
y <- numeric(length(sample_cols))
for (i in seq_along(sample_cols)) {
  y[i] <- as.numeric(pheno[oid_row, get(sample_cols[i])])
  
  # Progress reporting (less frequent to reduce output noise)
  if (i %% 1000 == 0 || i == length(sample_cols)) {
    cat("Processed", i, "of", length(sample_cols), "samples\n")
  }
}

#impute any missing protein level data (shouldnt be any)
na_count <- sum(is.na(y))
if (na_count > 0) {
  cat("Imputing", na_count, "missing phenotype values...\n")
  y[is.na(y)] <- mean(y, na.rm=TRUE)
}

#match samples genotype and protein data by sample ID
geno_samples <- genotypes$IID
common_samples <- intersect(geno_samples, sample_cols)

#use matched samples
if (length(common_samples) > 0) {
  cat("Found", length(common_samples), "samples in common between genotype and phenotype data\n")
  
  #find the indices of common samples in the genotype data
  geno_idx <- match(common_samples, geno_samples)
  #subset the genotype matrix to include only common samples
  X_matched <- X[geno_idx, , drop=FALSE]
  
  #find the indices of common samples in the phenotype data
  pheno_idx <- match(common_samples, sample_cols)
  #subset the protein level vector to include only common samples
  y_matched <- y[pheno_idx]
  
} else {
  #if no commonality found (this shouldnt occur)
  stop("Error: No common samples found between genotype and protein level data")
}

#run susieR
cat("Running SuSiE analysis...\n")

#store SNP names before running SuSiE
snp_names <- colnames(X_matched)

#check for issues
if (ncol(X_matched) == 0) {
  stop("Error: No SNPs to analyze after matching samples")
}

if (length(y_matched) == 0) {
  stop("Error: No phenotype data after matching samples")
}

if (length(y_matched) != nrow(X_matched)) {
  stop(paste("Error: Mismatched dimensions. X has", nrow(X_matched), 
             "rows but y has", length(y_matched), "elements"))
}

#run SuSiE
fitted_model <- tryCatch({
  susie(X_matched, y_matched, 
        L=10, 
        scaled_prior_variance=0.2,
        residual_variance=NULL, 
        prior_weights=NULL, 
        null_weight=NULL, # 0 is default
        standardize=TRUE, 
        intercept=TRUE,
        estimate_residual_variance=TRUE, 
        estimate_prior_variance=TRUE,
        estimate_prior_method='EM', # optim, EM, simple
        check_null_threshold=0, 
        prior_tol=1e-9,
        residual_variance_upperbound=Inf,
        coverage=0.95, 
        min_abs_corr=0.5,
        compute_univariate_zscore=FALSE,
        na.rm=FALSE, 
        max_iter=100, 
        tol=0.001,
        verbose=FALSE, 
        track_fit=FALSE)
}, error = function(e) {
  stop(paste("Error in SuSiE analysis:", e$message))
})

#save PIPs
snp_pips <- data.frame(
  SNP = snp_names,
  PIP = fitted_model$pip
)
snp_pips <- snp_pips[order(-snp_pips$PIP), ]
write.table(snp_pips, file = paste0(output_prefix, "_pips.tsv"), sep = "\t", quote = FALSE, row.names = FALSE)
cat("PIPs saved to", paste0(output_prefix, "_pips.tsv"), "\n")

#get credible sets with purity filtering
cs <- tryCatch({
  susie_get_cs(fitted_model, 
               X = X_matched,
               Xcorr=NULL,
               coverage = 0.95, 
               min_abs_corr = 0.5,
               dedup = TRUE,
               squared = FALSE)
}, error = function(e) {
  cat("Error getting credible sets:", conditionMessage(e), "\n")
  return(list())
})

#initialize empty dataframe
cs_results <- data.frame(
  SNP = character(),
  CredibleSet = integer(),
  PIP = numeric(),
  stringsAsFactors = FALSE
)

#check for credible set structure, then process
if (!is.null(cs$sets) && length(cs$sets) > 0) {
  #'sets' structure
  cat("Processing credible sets from 'sets' structure...\n")
  
  for (i in seq_along(cs$sets)) {
    indices <- cs$sets[[i]]
    temp_results <- data.frame(
      SNP = snp_names[indices],
      CredibleSet = i,
      PIP = fitted_model$pip[indices],
      stringsAsFactors = FALSE
    )
    cs_results <- rbind(cs_results, temp_results)
  }
} else {
  #find individual L sets (L1, L2)
  cat("Looking for individual L credible sets...\n")
  
  #check if directly in cs
  l_pattern <- "^L[0-9]+$"
  l_names <- names(cs)[grep(l_pattern, names(cs))]
  
  if (length(l_names) > 0) {
    for (l_name in l_names) {
      set_number <- as.integer(gsub("L", "", l_name))
      indices <- cs[[l_name]]
      if (length(indices) > 0) {
        temp_results <- data.frame(
          SNP = snp_names[indices],
          CredibleSet = set_number,
          PIP = fitted_model$pip[indices],
          stringsAsFactors = FALSE
        )
        cs_results <- rbind(cs_results, temp_results)
      }
    }
  } else if (!is.null(cs$cs)) {
    #if neither, check if cs$cs exists
    l_names <- names(cs$cs)[grep(l_pattern, names(cs$cs))]
    
    if (length(l_names) > 0) {
      for (l_name in l_names) {
        set_number <- as.integer(gsub("L", "", l_name))
        indices <- cs$cs[[l_name]]
        if (length(indices) > 0) {
          temp_results <- data.frame(
            SNP = snp_names[indices],
            CredibleSet = set_number,
            PIP = fitted_model$pip[indices],
            stringsAsFactors = FALSE
          )
          cs_results <- rbind(cs_results, temp_results)
        }
      }
    }
  }
}

#sort and save if we have results
if (nrow(cs_results) > 0) {
  #sort by credible set and PIP
  cs_results <- cs_results[order(cs_results$CredibleSet, -cs_results$PIP), ]
  
  #save file
  write.table(cs_results, 
              file = paste0(output_prefix, "_susie_cis_cs.tsv"),
              sep = "\t", 
              quote = FALSE, 
              row.names = FALSE)
  
  cat("Credible sets saved to", paste0(output_prefix, "_susie_cis_cs.tsv"), "\n")
} else {
  cat("No credible sets found in any format.\n")
}

#clean up
if (TRUE) {  #set to FALSE to keep temporary files
  cat("Cleaning up temporary files...\n")
  temp_files <- c(
    paste0(geno_output, ".raw"),
    paste0(geno_output, ".log"),
    paste0(geno_output, ".nosex")
  )
  for (file in temp_files) {
    if (file.exists(file)) {
      unlink(file)
    }
  }
}
```

**Run Cis SusieR in Parallel**
```{bash}
#!/bin/bash

#handle interrupt signals gracefully
trap 'echo "Script interrupted, exiting gracefully..."; exit 1' INT TERM

#input file containing OlinkID, chr, left, right data
INPUT_FILE="$HOME/ancestry_specific_folder/0.05cis_locations.tsv"

#Rscript path
R_SCRIPT="$HOME/ancestry_specific_folder/cis_susie.R"

#error log file
ERROR_LOG="$HOME/ancestry_specific_folder/susie_cis_analysis_errors.log"

#success log file
SUCCESS_LOG="$HOME/ancestry_specific_folder/susie_cis_analysis_complete.log"

#option to resume from a specific line number
START_LINE=2  #default to 2 to skip header

gunzip ~/ancestry_specific_folder/tensor_lm_table.bed.gz

#parse command line arguments
if [ $# -gt 0 ]; then
  case $1 in
    --start-from=*)
      START_LINE="${1#*=}"
      shift
      ;;
    *)
      echo "Unknown option: $1"
      echo "Usage: $0 [--start-from=LINE_NUMBER]"
      exit 1
      ;;
  esac
fi

#initialize or clear the error log
echo "SuSiE Analysis Error Log - $(date)" > "$ERROR_LOG"
echo "--------------------------------" >> "$ERROR_LOG"

#initialize or clear the success log
echo "SuSiE Analysis Success Log - $(date)" > "$SUCCESS_LOG"
echo "--------------------------------" >> "$SUCCESS_LOG"

#check if the input file exists
if [ ! -f "$INPUT_FILE" ]; then
    echo "Error: Input file not found: $INPUT_FILE"
    exit 1
fi

#count total number of analyses to run
TOTAL_ANALYSES=$(wc -l < "$INPUT_FILE")
TOTAL_ANALYSES=$((TOTAL_ANALYSES - 1))  #subtract 1 for header line
CURRENT_ANALYSIS=$((START_LINE - 2))    #adjust for header and 0-based counting

#create output directory 
OUTPUT_DIR="$HOME/ancestry_specific_folder/cis_susieR"
if [ ! -d "$OUTPUT_DIR" ]; then
    mkdir -p "$OUTPUT_DIR"
fi

#process each data line sequentially
#use head and tail to extract the relevant portion of the file
while read -r line; do
    IFS=$'\t' read -r oid chr left right <<< "$line"
    
    #skip if line is empty
    if [ -z "$oid" ]; then
        continue
    fi
    
    #extract the numerical part of the chromosome
    chr_num=$(echo "$chr" | grep -oP 'chr\K[0-9]+')
    
    #if chr_num is empty, try another approach (in case 'chr' prefix is missing)
    if [ -z "$chr_num" ]; then
        chr_num=$(echo "$chr" | grep -oP '[0-9]+')
    fi
    
    #skip if we couldn't extract a chromosome number
    if [ -z "$chr_num" ]; then
        echo "Error: Could not extract chromosome number from $chr"
        echo "Error: Could not extract chromosome number from $chr" >> "$ERROR_LOG"
        continue
    fi
    
    #update progress counter
    CURRENT_ANALYSIS=$((CURRENT_ANALYSIS + 1))
    echo "[$CURRENT_ANALYSIS/$TOTAL_ANALYSES] Processing $oid on $chr: positions $left-$right"
    
    #debug info about files
    echo "  - Checking phenotype file..."
    if [ -f "~/ancestry_specific_folder/tensor_lm_table.bed" ]; then
        file_size=$(ls -lh "~/ancestry_specific_folder/tensor_lm_table.bed" | awk '{print $5}')
        echo "  - Phenotype file exists (size: $file_size)"
    else
        echo "  - WARNING: Phenotype file does not exist!"
    fi
    
    #run the Rscript with a timeout
    output_file=$(mktemp)
    timeout 600 Rscript "$R_SCRIPT" --oid "$oid" --chr "$chr_num" 2>&1 | tee "$output_file"
    exit_code=$?
    
    #handle timeout or other errors
    if [ $exit_code -eq 124 ]; then
        echo "TIMEOUT: Analysis for $oid on chromosome $chr_num took too long and was terminated"
        echo "TIMEOUT: Analysis for $oid on chromosome $chr_num - $(date)" >> "$ERROR_LOG"
        echo "----------------------------------------" >> "$ERROR_LOG"
        echo "Analysis exceeded the 10-minute timeout limit." >> "$ERROR_LOG"
        echo "----------------------------------------" >> "$ERROR_LOG"
    elif [ $exit_code -ne 0 ]; then
        echo "ERROR processing $oid on chromosome $chr_num"
        #log the error
        echo "Error in $oid chromosome $chr_num (positions $left-$right) - $(date)" >> "$ERROR_LOG"
        echo "----------------------------------------" >> "$ERROR_LOG"
        cat "$output_file" >> "$ERROR_LOG"
        echo "----------------------------------------" >> "$ERROR_LOG"
    else
        #log success
        echo "Successfully processed $oid on chromosome $chr_num - $(date)" >> "$SUCCESS_LOG"
    fi
    
    #clean up temporary file
    rm "$output_file"
    
    echo "----------------------------------------"
done < <(tail -n +$START_LINE "$INPUT_FILE")

gzip ~/ancestry_specific_folder/tensor_lm_table.bed

echo "analyses complete"
```

**Combine Individual Cis SusieR Files**
```{r}
library(data.table)
library(dplyr)

#get all files
files <- list.files("~/ancestry_specific_folder/cis_susieR", pattern = "OID.*_susie_cis_cs\\.tsv$", full.names = TRUE)

#read first file to establish expected structure
first_file <- fread(files[1])
expected_cols <- colnames(first_file)
expected_ncols <- ncol(first_file)

print(paste("Expected structure:", expected_ncols, "columns"))
print(paste("Expected columns:", paste(expected_cols, collapse = ", ")))

#initialize list and counters
data_list <- list()
good_files <- 0
problem_oids <- c()

#check each file
for(i in 1:length(files)) {
  file_name <- basename(files[i])
  oid <- regmatches(file_name, regexpr("OID\\d+", file_name))
  
  tryCatch({
    data <- suppressWarnings(fread(files[i]))
    
    #check if structure matches
    if(ncol(data) != expected_ncols || !identical(colnames(data), expected_cols)) {
      problem_oids <- c(problem_oids, oid)
      print(paste("PROBLEM OID:", oid, "- has", ncol(data), "columns:", paste(colnames(data), collapse = ", ")))
    } else {
      #add OlinkID and store
      data$OlinkID <- oid
      data_list[[length(data_list) + 1]] <- data
      good_files <- good_files + 1
    }
    
  }, warning = function(w) {
    if(grepl("Stopped early", w$message)) {
      print(paste("CORRUPT DATA in OID:", oid, "- Warning:", w$message))
    }
    #still try to read the file
    data <- fread(files[i])
    data$OlinkID <- oid
    data_list[[length(data_list) + 1]] <- data
    good_files <<- good_files + 1
  }, error = function(e) {
    problem_oids <- c(problem_oids, oid)
    print(paste("ERROR OID:", oid, "- cannot read file:", e$message))
  })
  
  if(i %% 100 == 0) print(paste("Checked", i, "files"))
}

print(paste("Good files:", good_files))
print(paste("Problem files:", length(problem_oids)))

if(length(problem_oids) > 0) {
  print("All problem OIDs:")
  print(problem_oids)
}

#merge good files
if(length(data_list) > 0) {
  merged_data <- rbindlist(data_list)
  
  #save
  fwrite(merged_data, "~/ancestry_specific_folder/susieR/0.05cis_finemap.tsv", sep = "\t")
  print(paste("Merged", nrow(merged_data), "rows from", good_files, "files"))
  print("Saved to: ~/ancestry_specific_folder/susieR/0.05cis_finemap.tsv")
} else {
  print("No good files to merge!")
}
```

**Trans SusieR (~/ancestry_specific_folder/trans_susie.R)**
```{r}
library(susieR)
library(data.table)
library(argparse)

#setup command line argument parsing
parser <- ArgumentParser()
parser$add_argument("--oid", required=TRUE, help="OID number")
parser$add_argument("--chr", required=TRUE, help="Chromosome number")
args <- parser$parse_args()

#define file paths
genotype_prefix <- "~/ancestry_specific_folder/ancestry_final"
phenotype_file <- "~/ancestry_specific_folder/tensor_lm_table.bed"
positions_file <- paste0("~/ancestry_specific_folder/0.05trans/0.05phenotype_subsets/", args$oid, "_chr", args$chr, "_locations.tsv")
output_prefix <- paste0("~/ancestry_specific_folder/trans_susieR/", args$oid, "_chr", args$chr, "/", args$oid, "_chr", args$chr)
output_dir <- paste0("~/ancestry_specific_folder/trans_susieR/", args$oid, "_chr", args$chr, "/")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

#read in the position file
positions <- fread(positions_file)
if(ncol(positions) == 4) {
  # File has gene_id column - remove it
  positions <- positions[, 2:4]
}

#rename columns
colnames(positions) <- c("phenotype_id", "chr", "pos")

#extract region of interest based on the file's positions
min_pos <- min(positions$pos) - 1000000
max_pos <- max(positions$pos) + 1000000
target_positions <- positions$pos

#run PLINK2 command to make raw file of this region
geno_output <- paste0(output_prefix, "_genotypes")
cmd <- sprintf("plink2 --pfile %s --chr %s --from-bp %d --to-bp %d --export A --out %s", 
               genotype_prefix, args$chr, min_pos, max_pos, geno_output)
system(cmd)

#check if file exists and read in extracted genotypes
geno_file <- paste0(geno_output, ".raw")
if (!file.exists(geno_file)) {
  stop("PLINK2 extraction failed. Check PLINK log file for details.")
}

genotypes <- fread(geno_file)

#remove metadata
meta_cols <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")
meta_cols <- meta_cols[meta_cols %in% colnames(genotypes)]
geno_cols <- setdiff(colnames(genotypes), meta_cols)

#find chr and pos of each SNP IDs; this is comprehensive to many SNP ID formats
extract_pos <- function(snp_id) {
  #if the SNP ID contains colons
  if (grepl(":", snp_id)) { 
    parts <- strsplit(snp_id, ":")[[1]]
    pos_part <- parts[2]
    #if the SNP ID contains underscores and colons
    if (grepl("_", pos_part)) {
      pos <- as.numeric(strsplit(pos_part, "_")[[1]][1])
    } else {
      pos <- as.numeric(pos_part)
    }
    return(pos)
    #if the SNP ID contains underscores only
  } else if (grepl("_", snp_id)) {
    parts <- strsplit(snp_id, "_")[[1]]
    if (length(parts) >= 2) {
      return(as.numeric(parts[2]))
    }
  }
  return(NA) #if the position cannot be found
}

snp_positions <- sapply(geno_cols, extract_pos) #identify the positions
matching_snps <- geno_cols #use all snps in window

#create genotype matrix
X <- matrix(as.numeric(as.matrix(genotypes[, matching_snps, with=FALSE])), 
            nrow=nrow(genotypes), ncol=length(matching_snps))
colnames(X) <- matching_snps

#impute missing genotype data (shouldnt be any)
na_count <- sum(is.na(X))
if (na_count > 0) {
  for (j in 1:ncol(X)) {
    na_idx <- is.na(X[,j])
    if (any(na_idx)) {
      X[na_idx, j] <- mean(X[!na_idx, j])
    }
  }
}

#read in protein level data
pheno <- fread(phenotype_file, header=TRUE)
oid_rows <- which(pheno$gene_id == args$oid)
if (length(oid_rows) == 0) {
  stop(paste("Gene", args$oid, "not found in phenotype file"))
}
oid_row <- oid_rows[1]

#remove any metadata
meta_cols <- c("#chr", "chr", "start", "end", "gene_id")
meta_cols <- meta_cols[meta_cols %in% colnames(pheno)]
sample_cols <- setdiff(colnames(pheno), meta_cols)

#find the relevant protein level data
y <- numeric(length(sample_cols))
for (i in seq_along(sample_cols)) {
  y[i] <- as.numeric(pheno[oid_row, get(sample_cols[i])])
  
  #progress
  if (i %% 500 == 0 || i == length(sample_cols)) {
    cat("Processed", i, "of", length(sample_cols), "samples\n")
  }
}

#impute any missing protein level data (shouldnt be any)
na_count <- sum(is.na(y))
if (na_count > 0) {
  y[is.na(y)] <- mean(y, na.rm=TRUE)
}

#match samples genotype and protein data by sample ID
geno_samples <- genotypes$IID
common_samples <- intersect(geno_samples, sample_cols)

#use matched samples, if found
if (length(common_samples) > 0) {
  #find the indices of common samples in the genotype data
  geno_idx <- match(common_samples, geno_samples)
  #subset the genotype matrix to include only common samples (should be like this already)
  X_matched <- X[geno_idx, , drop=FALSE]
  
  #find the indices of common samples in the phenotype data
  pheno_idx <- match(common_samples, sample_cols)
  #subset the protein level vector to include only common samples (should be like this already)
  y_matched <- y[pheno_idx]
  
} else {
  #if no commonality found, this shouldnt occur
  stop("Error: No common samples found between genotype and protein level data")
}

#run susieR
cat("Running SuSiE analysis...\n")

#store SNP names before running SuSiE
snp_names <- colnames(X_matched)

#these are the same parameters tensorqtl used
fitted_model <- susie(X_matched, y_matched, 
                      L=10, 
                      scaled_prior_variance=0.2,
                      residual_variance=NULL, 
                      prior_weights=NULL, 
                      null_weight=NULL, #0 is default
                      standardize=TRUE, 
                      intercept=TRUE,
                      estimate_residual_variance=TRUE, 
                      estimate_prior_variance=TRUE,
                      estimate_prior_method='EM', #optim, EM, simple
                      check_null_threshold=0, 
                      prior_tol=1e-9,
                      residual_variance_upperbound=Inf,
                      coverage=0.95, 
                      min_abs_corr=0.5,
                      compute_univariate_zscore=FALSE,
                      na.rm=FALSE, 
                      max_iter=100, 
                      tol=0.001,
                      verbose=FALSE, 
                      track_fit=FALSE)

#save PIPs directly to TSV using the stored SNP names
snp_pips <- data.frame(
  SNP = snp_names,
  PIP = fitted_model$pip
)
snp_pips <- snp_pips[order(-snp_pips$PIP), ]

#get credible sets with 0.5 purity filtering
cs <- tryCatch({
  susie_get_cs(fitted_model, 
               X = X_matched,
               Xcorr=NULL,
               coverage = 0.95, 
               min_abs_corr = 0.5,
               dedup = TRUE,
               squared = FALSE)
}, error = function(e) {
  cat("Error getting credible sets:", conditionMessage(e), "\n")
  return(list())
})

#initialize empty results dataframe
cs_results <- data.frame(
  SNP = character(),
  CredibleSet = integer(),
  PIP = numeric(),
  stringsAsFactors = FALSE
)

#check which credible set structure we have and process accordingly
if (!is.null(cs$sets) && length(cs$sets) > 0) {
  #'sets' structure
  cat("Processing credible sets from 'sets' structure...\n")
  
  for (i in seq_along(cs$sets)) {
    indices <- cs$sets[[i]]
    temp_results <- data.frame(
      SNP = snp_names[indices],
      CredibleSet = i,
      PIP = fitted_model$pip[indices],
      stringsAsFactors = FALSE
    )
    cs_results <- rbind(cs_results, temp_results)
  }
} else {
  #try to find individual L sets (L1, L2, etc.)
  cat("Looking for individual L credible sets...\n")
  
  #check if they're directly in cs
  l_pattern <- "^L[0-9]+$"
  l_names <- names(cs)[grep(l_pattern, names(cs))]
  
  if (length(l_names) > 0) {
    for (l_name in l_names) {
      set_number <- as.integer(gsub("L", "", l_name))
      indices <- cs[[l_name]]
      if (length(indices) > 0) {
        temp_results <- data.frame(
          SNP = snp_names[indices],
          CredibleSet = set_number,
          PIP = fitted_model$pip[indices],
          stringsAsFactors = FALSE
        )
        cs_results <- rbind(cs_results, temp_results)
      }
    }
  } else if (!is.null(cs$cs)) {
    #if neither of the above, check if cs$cs exists
    l_names <- names(cs$cs)[grep(l_pattern, names(cs$cs))]
    
    if (length(l_names) > 0) {
      for (l_name in l_names) {
        set_number <- as.integer(gsub("L", "", l_name))
        indices <- cs$cs[[l_name]]
        if (length(indices) > 0) {
          temp_results <- data.frame(
            SNP = snp_names[indices],
            CredibleSet = set_number,
            PIP = fitted_model$pip[indices],
            stringsAsFactors = FALSE
          )
          cs_results <- rbind(cs_results, temp_results)
        }
      }
    }
  }
}

#sort and save if we have results
if (nrow(cs_results) > 0) {
  #sort by credible set and PIP
  cs_results <- cs_results[order(cs_results$CredibleSet, -cs_results$PIP), ]
  
  #save
  write.table(cs_results, 
              file = paste0(output_prefix, "_susie_trans_cs.tsv"),
              sep = "\t", 
              quote = FALSE, 
              row.names = FALSE)
  
  cat("Credible sets saved to", paste0(output_prefix, "_credible_sets.tsv"), "\n")
} else {
  cat("No credible sets found in any format.\n")
}

#clean up temp files
if (TRUE) {  # Set to FALSE to keep temporary files
  cat("Cleaning up temporary files...\n")
  temp_files <- c(
    paste0(geno_output, ".raw"),
    paste0(geno_output, ".log"),
    paste0(geno_output, ".nosex")
  )
  for (file in temp_files) {
    if (file.exists(file)) {
      unlink(file)
    }
  }
}
```

**Run Trans SusieR in Parallel**
```{bash}
#!/bin/bash

#directory containing the location files
LOCATIONS_DIR="$HOME/ancestry_specific_folder/0.05trans/0.05phenotype_subsets"

#Rscript path
R_SCRIPT="$HOME/ancestry_specific_folder/trans_susie.R"

#error log file
ERROR_LOG="$HOME/ancestry_specific_folder/susieR/susie_trans_analysis_errors.log"

#initialize or clear the error log
echo "SuSiE Analysis Error Log - $(date)" > "$ERROR_LOG"
echo "--------------------------------" >> "$ERROR_LOG"

#loop through each location file
for location_file in "$LOCATIONS_DIR"/OID*_chr*_locations.tsv; do
    #extract the OID number and chromosome number from the filename
    filename=$(basename "$location_file")
    oid_num=$(echo "$filename" | grep -oP 'OID\K\d+')
    chr_num=$(echo "$filename" | grep -oP 'chr\K[0-9]+')

    echo "Processing $filename: OID$oid_num on chromosome $chr_num"

    #run the R script with the appropriate OID and chromosome as arguments
    #use tee to show output while also capturing it
    output_file=$(mktemp)
    set -o pipefail
    Rscript "$R_SCRIPT" --oid "OID$oid_num" --chr "$chr_num" 2>&1 | tee "$output_file"
    exit_code=$?

    #check if the command succeeded
    if [ $exit_code -ne 0 ]; then
        echo "ERROR processing OID$oid_num chromosome $chr_num"
        #log the error
        echo "Error in OID$oid_num chromosome $chr_num - $(date)" >> "$ERROR_LOG"
        echo "----------------------------------------" >> "$ERROR_LOG"
        cat "$output_file" >> "$ERROR_LOG"
        echo "----------------------------------------" >> "$ERROR_LOG"
    fi

    #clean up temp file
    rm "$output_file"

    echo "----------------------------------------"
done
```

**Combine Individual Trans SusieR Files**
```{r}
library(data.table)
library(dplyr)

#get all trans SuSiE files from the directory structure
base_dir <- "~/ancestry_specific_folder/trans_susieR"
files <- list.files(base_dir, pattern = ".*_susie_trans_cs\\.tsv$", full.names = TRUE, recursive = TRUE)

#read first file to establish expected structure
first_file <- fread(files[1])
expected_cols <- colnames(first_file)
expected_ncols <- ncol(first_file)

print(paste("Expected structure:", expected_ncols, "columns"))
print(paste("Expected columns:", paste(expected_cols, collapse = ", ")))

#initialize list and counters
data_list <- list()
good_files <- 0
problem_oids <- c()

#check each file
for(i in 1:length(files)) {
  file_path <- files[i]
  file_name <- basename(file_path)
  
  #extract OID and chromosome from filename
  OlinkID <- regmatches(file_name, regexpr("OID\\d+", file_name))
  if(length(OlinkID) == 0) {
    # Try to extract just OID if the pattern is different
    OlinkID <- regmatches(file_name, regexpr("OID\\d+", file_name))
  }
  
  tryCatch({
    data <- suppressWarnings(fread(file_path))
    
    #skip empty files
    if(nrow(data) == 0) {
      print(paste("EMPTY FILE:", OlinkID, "- skipping"))
      next
    }
    
    #check if structure matches
    if(ncol(data) != expected_ncols || !identical(colnames(data), expected_cols)) {
      problem_oids <- c(problem_oids, OlinkID)
      print(paste("PROBLEM OID:", OlinkID, "- has", ncol(data), "columns:", paste(colnames(data), collapse = ", ")))
    } else {
      #add OlinkID_chr identifier and store
      data$OlinkID <- OlinkID
      data_list[[length(data_list) + 1]] <- data
      good_files <- good_files + 1
    }
    
  }, warning = function(w) {
    if(grepl("Stopped early", w$message)) {
      print(paste("CORRUPT DATA in OID:", OlinkID, "- Warning:", w$message))
    }
    #still try to read the file
    tryCatch({
      data <- fread(file_path)
      if(nrow(data) > 0) {
        data$OlinkID_chr <- OlinkID
        data_list[[length(data_list) + 1]] <- data
        good_files <<- good_files + 1
      }
    }, error = function(e2) {
      problem_oids <<- c(problem_oids, OlinkID)
      print(paste("ERROR after warning for OID:", OlinkID, "- cannot read file:", e2$message))
    })
  }, error = function(e) {
    problem_oids <<- c(problem_oids, OlinkID)
    print(paste("ERROR OID:", OlinkID, "- cannot read file:", e$message))
  })
  
  if(i %% 100 == 0) print(paste("Checked", i, "files"))
}

print(paste("Good files:", good_files))
print(paste("Problem files:", length(problem_oids)))

if(length(problem_oids) > 0) {
  print("All problem OIDs:")
  print(problem_oids)
}

#merge good files
if(length(data_list) > 0) {
  merged_data <- rbindlist(data_list)
  
  #save to the parent directory
  output_file <- "~/ancestry_specific_folder/susieR/0.05trans_finemap.tsv"
  fwrite(merged_data, output_file, sep = "\t")
  print(paste("Merged", nrow(merged_data), "rows from", good_files, "files"))
  print(paste("Saved to:", output_file))
  
  #print summary statistics
  print(paste("Unique OIDs processed:", length(unique(merged_data$OlinkID_chr))))
  
} else {
  print("No good files to merge!")
}
```

**Filter Final File**
- Group by OID_#, credible set label, and chromosome to isolate credible sets
- Keep only credible sets with greater than or equal to 0.9 coverage
- Relabel credible set labels to make each unique within OID group
- Remove _# after OID#####
```{r}
library(dplyr)
library(data.table)
library(stringr)

trans_condensed <- fread("~/ancestry_specific_folder/susieR/0.05trans_finemap.tsv", header=TRUE)
colnames(trans_condensed) <- c("variant_id", "cs_label", "pip", "gene_id")

#add chr column
trans_condensed$chr <- str_extract(trans_condensed$variant_id, "chr\\d+")
trans_condensed$chr[is.na(trans_condensed$chr)] <- "chr23"
trans_condensed$base_gene_id <- str_extract(trans_condensed$gene_id, "^[^_]+")

#keep only credible sets that add to .9 (conservative filter)
trans_filtered <- trans_condensed %>%
  group_by(gene_id, cs_label, chr) %>%
  filter(sum(pip) >= 0.9) %>%
  ungroup() %>%
  mutate(
    #create a unique identifier for each credible set
    cs_combo = paste(gene_id, cs_label, chr, sep = "_"),
    #assign ID based on first appearance order
    cs_id = as.numeric(factor(cs_combo, levels = unique(cs_combo)))
  ) %>%
  select(-cs_combo)

#relabel credible sets to make unique cs_label per protein
trans_cs_relabel <- trans_filtered %>%
  mutate(original_order = row_number()) %>%
  group_by(base_gene_id) %>%
  mutate(
    #assign sequential L numbers based on unique cs_id values
    new_cs_label = paste0("L", dense_rank(cs_id))
  ) %>%
  ungroup() %>%
  arrange(original_order) %>%
  select(-cs_label, -original_order) %>%
  rename(cs_label = new_cs_label)

#rearrange
trans_reorder <- trans_cs_relabel %>%
  select(base_gene_id, cs_label, variant_id, pip)
colnames(trans_reorder) <- c("OlinkID", "CredibleSet", "SNP", "PIP")

#save
write.table(trans_reorder, "~/ancestry_specific_folder/susieR/0.05trans_finemap_edited.tsv", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
```
