---
title: "Replicate Statistically Significant MESA SNPs in UKB"
author: "Claudia Krueger"
date: "2025-07-10"
output: html_document
---

**Find Cis Qvalues to Isolate Statistically Significant MESA cis-SNPs**
TensorQTL only outputs a single qvalue/protein, therefore SNP level qvalues need to be calculated with the qvalue R package

```{r}
library(data.table)
library(dplyr)
library(qvalue)
set.seed(123)

#initiate list
cis_data <- list()
pvals <- list()

#loop through chromosomes to save to individual variables
for(chr in 1:23) {
  cat("Starting Chr:", chr, "\n")
  gz_con <- gzfile(paste0("~/ancestry_specific_folder/tensorqtl/nominal/nominalchr", chr, "cis_qtls.gz"), "rt")
  cis_data[[chr]] <- read.table(gz_con, header = TRUE, sep = "\t")
  close(gz_con)
  pvals[[chr]] <- cis_data[[chr]]$pval_nominal
}

#combine all p-values and track which chromosome each came from
all_pvals <- unlist(pvals)
chr_indices <- rep(1:23, sapply(pvals, length))

#calculate q-values
qobj <- qvalue(all_pvals)
all_qvals <- qobj$qvalues

#split q-values back to original chromosome structure
qvals_by_chr <- split(all_qvals, chr_indices)

#add q-values back to each chromosome's data and filter
filtered_cis_data <- list()

for(chr in 1:23) {
  cis_data[[chr]]$qval <- qvals_by_chr[[chr]] #add q-values to this chromosome's data
  filtered_cis_data[[chr]] <- cis_data[[chr]][cis_data[[chr]]$qval < 0.05, ] #filter for significant q-values (< 0.05)
  cat("Chr", chr, "- Total:", nrow(cis_data[[chr]]), "Significant:", nrow(filtered_cis_data[[chr]]), "\n")
}

#combine all filtered data into one dataframe
all_significant_qtls <- do.call(rbind, filtered_cis_data)

#write out the filtered combined file
write.table(all_significant_qtls, "~/ancestry_specific_folder/tensorqtl/nominal/all_cis_qtls_qval_package_filtered.tsv", sep = "\t", row.names = FALSE, quote = FALSE)

#print summary from qvalue object
cat("Estimated pi0 (proportion of true nulls):", qobj$pi0, "\n")
cat("Total significant QTLs (q < 0.05):", nrow(all_significant_qtls), "\n")
```

**Isolate Statistically Significant MESA trans-SNPs**
```{bash}
#unzip original TensorQTL output
gunzip ~/ancestry_specific_folder/tensorqtl/trans/trans_perm.tsv.gz 

#filter by qval column < 0.05
awk 'NR==1 || $14 < 0.05' ~/ancestry_specific_folder/tensorqtl/trans/trans_perm.tsv > ~/ancestry_specific_folder/tensorqtl/trans/qval_filt_trans_perm.tsv &
```

**Validate MESA cis- and trans-pQTL SNPs in UKBB Olink pQTL summary statistics**
```{python}
import pandas as pd
import numpy as np
import os
import gzip
import re
from pathlib import Path
import subprocess

def extract_oid_from_filename(filename):
    #extract OID number from filename
    match = re.search(r'OID(\d+(?:\.\d+)?)', filename)
    return match.group(1) if match else None

def load_cis_qtl_data(cis_qtl_file):
    #load cis QTL data and parse genomic coordinates
    print("Loading cis QTL data...")
    cis_qtl = pd.read_csv(cis_qtl_file, sep='\t')

    variant_parts = cis_qtl['variant_id'].str.split(':', expand=True)
    cis_qtl['CHROM'] = variant_parts[0]
    cis_qtl['GENPOS'] = variant_parts[1].astype(int)
    cis_qtl['QTL_REF'] = variant_parts[2]
    cis_qtl['QTL_ALT'] = variant_parts[3]

    cis_qtl['allele_key'] = cis_qtl['QTL_REF'] + ':' + cis_qtl['QTL_ALT']

    cis_qtl = cis_qtl.rename(columns={
        'phenotype_id': 'OlinkID',
        'pval_nominal': 'QTL_Pvalue',
        'slope': 'QTL_BETA',
        'slope_se': 'QTL_SE',
        'qval': 'QTL_qvalue',
        'af': 'QTL_AF'
    })

    cis_qtl['CHROM'] = cis_qtl['CHROM'].replace({'chrX': 'chr23', 'X': 'chr23', '23': 'chr23'})

    print(f"Loaded {len(cis_qtl)} cis QTL associations")
    return cis_qtl

def load_trans_qtl_data(trans_qtl_file):
    #oad trans QTL data and parse genomic coordinates
    print("Loading trans QTL data...")
    trans_qtl = pd.read_csv(trans_qtl_file, sep='\t')

    variant_parts = trans_qtl['variant_id'].str.split(':', expand=True)
    trans_qtl['CHROM'] = variant_parts[0]
    trans_qtl['GENPOS'] = variant_parts[1].astype(int)
    trans_qtl['QTL_REF'] = variant_parts[2]
    trans_qtl['QTL_ALT'] = variant_parts[3]

    trans_qtl['allele_key'] = trans_qtl['QTL_REF'] + ':' + trans_qtl['QTL_ALT']

    trans_qtl = trans_qtl.rename(columns={
        'phenotype_id': 'OlinkID',
        'pval': 'QTL_Pvalue',
        'b': 'QTL_BETA',
        'b_se': 'QTL_SE',
        'qval': 'QTL_qvalue',
        'af': 'QTL_AF'
    })

    trans_qtl['CHROM'] = trans_qtl['CHROM'].replace({'chrX': 'chr23', 'X': 'chr23', '23': 'chr23'})

    print(f"Loaded {len(trans_qtl)} trans QTL associations")
    return trans_qtl

def process_single_gwas_file(file_path, cis_qtl_data, trans_qtl_data, output_dir=None):
    #process a single GWAS file and create overlapping cis and trans files
    filename = file_path.name
    print(f"Processing: {filename}")

    oid = extract_oid_from_filename(filename)
    if not oid:
        print(f"Warning: Could not extract OID from {filename}")
        return None

    try:
        with gzip.open(file_path, 'rt') as f:
            df = pd.read_csv(f, sep='\t')

        all_rows = df.copy()

        if len(all_rows) == 0:
            print(f"No data in {filename}")
            return {
                'filename': filename,
                'oid': oid,
                'total_variants': 0,
                'cis_overlap': 0,
                'trans_overlap': 0,
                'cis_output_file': None,
                'trans_output_file': None
            }

        full_oid = f"OID{oid}"
        all_rows['OlinkID'] = full_oid
        all_rows['allele_key'] = (
            all_rows['ID'].str.split(':').str[2] + ':' +
            all_rows['ID'].str.split(':').str[3]
        )

        #rename UKBB columns once
        ukbb = all_rows.rename(columns={
            'ID': 'UKBB_ID',
            'ALLELE0': 'UKBB_A0',
            'ALLELE1': 'UKBB_A1',
            'N': 'UKBB_N',
            'BETA': 'UKBB_BETA',
            'SE': 'UKBB_SE',
            'Pvalue': 'UKBB_Pvalue'
        })

        #cis overlap
        cis_overlap = pd.DataFrame()
        qtl_cis_protein = cis_qtl_data[cis_qtl_data['OlinkID'] == full_oid]
        if len(qtl_cis_protein) > 0:
            cis_overlap = pd.merge(
                ukbb, qtl_cis_protein,
                on=['OlinkID', 'CHROM', 'GENPOS', 'allele_key'],
                how='inner'
            )
            cis_overlap['cis_trans'] = 'cis'
            cis_cols = ['OlinkID', 'CHROM', 'GENPOS', 'variant_id', 'UKBB_ID',
                        'QTL_Pvalue', 'UKBB_Pvalue', 'QTL_BETA', 'QTL_SE',
                        'UKBB_BETA', 'UKBB_SE', 'QTL_qvalue', 'QTL_REF',
                        'QTL_ALT', 'UKBB_A0', 'UKBB_A1', 'A1FREQ', 'UKBB_N',
                        'QTL_AF', 'cis_trans']
            cis_overlap = cis_overlap[[c for c in cis_cols if c in cis_overlap.columns]]

        #trans overlap
        trans_overlap = pd.DataFrame()
        qtl_trans_protein = trans_qtl_data[trans_qtl_data['OlinkID'] == full_oid]
        if len(qtl_trans_protein) > 0:
            trans_overlap = pd.merge(
                ukbb, qtl_trans_protein,
                on=['OlinkID', 'CHROM', 'GENPOS', 'allele_key'],
                how='inner'
            )
            trans_overlap['cis_trans'] = 'trans'
            trans_cols = ['OlinkID', 'CHROM', 'GENPOS', 'variant_id', 'UKBB_ID',
                          'QTL_Pvalue', 'UKBB_Pvalue', 'QTL_BETA', 'QTL_SE',
                          'UKBB_BETA', 'UKBB_SE', 'QTL_qvalue', 'QTL_REF',
                          'QTL_ALT', 'UKBB_A0', 'UKBB_A1', 'A1FREQ', 'UKBB_N',
                          'QTL_AF', 'cis_trans']
            trans_overlap = trans_overlap[[c for c in trans_cols if c in trans_overlap.columns]]

        #save
        cis_output_file = None
        trans_output_file = None

        if output_dir:
            for overlap_df, label in [(cis_overlap, 'cis'), (trans_overlap, 'trans')]:
                if len(overlap_df) > 0:
                    out_path = Path(output_dir) / f"{full_oid}_{label}_overlap.tsv"
                    overlap_df.to_csv(out_path, sep='\t', index=False)
                    print(f"Saved {len(overlap_df)} {label} overlapping variants to {out_path}")
                    try:
                        subprocess.run(['gzip', '-f', str(out_path)], check=True)
                        out_path = Path(str(out_path) + '.gz')
                        print(f"Compressed to {out_path}")
                    except subprocess.CalledProcessError as e:
                        print(f"Warning: Failed to gzip {out_path}: {e}")
                    if label == 'cis':
                        cis_output_file = str(out_path)
                    else:
                        trans_output_file = str(out_path)
                else:
                    print(f"No {label} overlapping variants found for {full_oid}")

        print(f"  Total variants: {len(all_rows)}")
        print(f"  Cis QTL variants for protein: {len(qtl_cis_protein)}  |  overlap: {len(cis_overlap)}")
        print(f"  Trans QTL variants for protein: {len(qtl_trans_protein)}  |  overlap: {len(trans_overlap)}")

        return {
            'filename': filename,
            'oid': oid,
            'full_oid': full_oid,
            'total_variants': len(all_rows),
            'cis_overlap': len(cis_overlap),
            'trans_overlap': len(trans_overlap),
            'cis_output_file': cis_output_file,
            'trans_output_file': trans_output_file
        }

    except Exception as e:
        print(f"Error processing {filename}: {str(e)}")
        return None

def process_gwas_files(folder_path, cis_qtl_file, trans_qtl_file, output_dir=None):
    """Process all GWAS files in folder against cis and trans QTL data"""
    if output_dir:
        Path(output_dir).mkdir(parents=True, exist_ok=True)
        print(f"Output directory: {output_dir}")

    cis_qtl_data = load_cis_qtl_data(cis_qtl_file)
    trans_qtl_data = load_trans_qtl_data(trans_qtl_file)

    folder = Path(folder_path)
    gwas_files = list(folder.glob("*.gz"))
    print(f"Found {len(gwas_files)} files to process...")

    summary_results = []

    for file_path in gwas_files:
        result = process_single_gwas_file(
            file_path, cis_qtl_data, trans_qtl_data, output_dir
        )
        if result:
            summary_results.append(result)
        print("---")

    if summary_results:
        summary_df = pd.DataFrame(summary_results)
        print(f"\n=== OVERALL SUMMARY ===")
        print(f"Files processed: {len(summary_results)}")
        print(f"Total variants: {summary_df['total_variants'].sum()}")
        print(f"Total cis overlap: {summary_df['cis_overlap'].sum()}")
        print(f"Total trans overlap: {summary_df['trans_overlap'].sum()}")

        if output_dir:
            summary_file = Path(output_dir) / "processing_summary.tsv"
            summary_df.to_csv(summary_file, sep='\t', index=False)
            print(f"Summary saved to: {summary_file}")

        return summary_df
    else:
        print("No files were successfully processed")
        return pd.DataFrame()

if __name__ == "__main__":
    folder_path = "~/ukbio/mesa_comparison/ancestry_specific_folder/"
    cis_qtl_file = "~/ancestry_specific_folder/tensorqtl/nominal/all_cis_qtls_qval_package_filtered.tsv"
    trans_qtl_file = "~/ancestry_specific_folder/tensorqtl/trans/trans_perm0.05.tsv"
    output_dir = "~/ukbio/mesa_comparison/ancestry_specific_folder/"

    summary = process_gwas_files(folder_path, cis_qtl_file, trans_qtl_file, output_dir=output_dir)

    if not summary.empty:
        print("\nFirst 5 rows of summary:")
        print(summary.head())
```

**Merge all overlapping cis SNPs**
```{bash}
zcat ~/ukbio/mesa_comparison/ancestry_specific_folder/OID*_cis_overlap.tsv.gz | head -1 > ~/ukbio/mesa_comparison/ancestry_specific_folder/00combined_cis_overlap.tsv

for file in ~/ukbio/mesa_comparison/ancestry_specific_folder/*_cis_overlap.tsv.gz; do
    zcat "$file" | tail -n +2 >> ~/ukbio/mesa_comparison/ancestry_specific_folder/00combined_cis_overlap.tsv
done
```

**Merge all overlapping trans SNPs**
```{bash}
zcat ~/ukbio/mesa_comparison/ancestry_specific_folder/OID*_trans_overlap.tsv.gz | head -1 > ~/ukbio/mesa_comparison/ancestry_specific_folder/00combined_trans_overlap.tsv

for file in zcat ~/ukbio/mesa_comparison/ancestry_specific_folder/*_trans_overlap.tsv.gz; do
    zcat "$file" | tail -n +2 >> zcat ~/ukbio/mesa_comparison/ancestry_specific_folder/00combined_trans_overlap.tsv
done
```

**Calculate pi1**
```{r}
library(data.table)
library(dplyr)
library(qvalue)
set.seed(123)

cis_data <- fread("~/ukbio/mesa_comparison/ancestry_specific_folder/00combined_cis_overlap.tsv")
cis_data$QTL_Pvalue[cis_data$QTL_Pvalue == 0] <- 1e-300
cis_data$UKBB_Pvalue[cis_data$UKBB_Pvalue == 0] <- 1e-300

trans_data <- fread("~/ukbio/mesa_comparison/ancestry_specific_folder/00combined_trans_overlap.tsv")
trans_data$QTL_Pvalue[trans_data$QTL_Pvalue == 0] <- 1e-300
trans_data$UKBB_Pvalue[trans_data$UKBB_Pvalue == 0] <- 1e-300

#extract p-values
cis_pvals_tensor <- cis_data$QTL_Pvalue
cis_pvals_ukbb <- cis_data$UKBB_Pvalue
trans_pvals_tensor <- trans_data$QTL_Pvalue
trans_pvals_ukbb <- trans_data$UKBB_Pvalue

#MESA/tensor qvalue calculations with adjusted lambda (extremely skewed, but we dont use this anyways)
cis_qobj_tensor <- qvalue(c(cis_pvals_tensor, 1))
trans_qobj_tensor <- qvalue(c(trans_pvals_tensor, 1))

#UKBB qvalue calculations with default parameters
cis_qobj_ukbb <- qvalue(cis_pvals_ukbb)
trans_qobj_ukbb <- qvalue(trans_pvals_ukbb)

#calculate π1 statistics
cis_pi1_tensor <- 1 - cis_qobj_tensor$pi0
cis_pi1_ukbb <- 1 - cis_qobj_ukbb$pi0
trans_pi1_tensor <- 1 - trans_qobj_tensor$pi0
trans_pi1_ukbb <- 1 - trans_qobj_ukbb$pi0

#calculate overlap statistics
cis_oid_overlap <- n_distinct(cis_data$OlinkID)
trans_oid_overlap <- n_distinct(trans_data$OlinkID)
cis_snp_overlap <- n_distinct(cis_data$variant_id)
trans_snp_overlap <- n_distinct(trans_data$variant_id)
cis_oid_snp_overlap <- nrow(cis_data)
trans_oid_snp_overlap <- nrow(trans_data)

#create results data frame
results <- data.frame(
  Dataset = c("MESA", "UKBB", "MESA", "UKBB"),
  Type = c("Cis", "Cis", "Trans", "Trans"),
  OID_Overlap = c(cis_oid_overlap, cis_oid_overlap, trans_oid_overlap, trans_oid_overlap),
  Unique_SNP_Overlap = c(cis_snp_overlap, cis_snp_overlap, trans_snp_overlap, trans_snp_overlap),
  OID_SNP_Pair_Overlap = c(cis_oid_snp_overlap, cis_oid_snp_overlap, trans_oid_snp_overlap, trans_oid_snp_overlap),
  pi1 = c(cis_pi1_tensor, cis_pi1_ukbb, trans_pi1_tensor, trans_pi1_ukbb),
  stringsAsFactors = FALSE
)

print(results)

#save
write.table(results, "~/ukbio/mesa_comparison/ancestry_specific_folder/00overlap_stats.tsv",
            sep = "\t", row.names = FALSE, quote = FALSE, col.names = TRUE)
```
