---
title: "TensorQTL Cis and Trans Susie Fine-Mapping"
author: "Claudia Krueger"
date: "2025-07-01"
output: html_document
---

**Create Dummy Covariate File**
```{python}
import pandas as pd
import numpy as np

#get fids from plink file
pheno_file = "~/ancestry_specific_folder/tensor_lm_table.bed.gz"
with pd.read_csv(pheno_file, sep='\t', iterator=True, chunksize=1) as reader:
    header = next(reader)
    sample_ids = header.columns[4:].tolist()

#create column of 1s
dummy_cov = pd.DataFrame({'intercept': np.ones(len(sample_ids))}, index=sample_ids)

#save covariates; transposed
dummy_cov.T.to_csv("~/ancestry_specific_folder/dummy_covariates.txt", sep='\t')
```

**Create Cis Location File**
```{r}
library(data.table)
library(dplyr)

data <- fread("~/ancestry_specific_folder/empirical/0.05cis_qtls.tsv", header=TRUE)

#filter by qvalue phenotype ids
oids <- data %>% select(phenotype_id)
locations <- fread("~/protein_locations/allchr_protein_location_file")
cis_locations <- left_join(oids, locations, by=c("phenotype_id" = "OlinkID"))

#save
write.table(cis_locations, "~/ancestry_specific_folder/0.05cis_locations.tsv", row.names = FALSE, quote=FALSE, col.names=TRUE, sep="\t")
```

**Cis Fine-Mapping**
```{bash}
python3 -m tensorqtl ~/ancestry_specific_folder/ancestry_final ~/ancestry_specific_folder/tensor_lm_table.bed.gz ~/ancestry_specific_folder/tensor_susie/0.05cis  --cis_output ~/ancestry_specific_folder/tensorqtl/empirical/0.05cis_qtls.tsv --covariates ~/ancestry_specific_folder/dummy_covariates.txt --mode cis_susie --seed 123
```

**Unpack Cis Susie Pickle**
```{python}
import pandas as pd
import pickle

#load the pickle file
with open("~/ancestry_specific_folder/tensor_susie/0.05cis.SuSiE.pickle", 'rb') as f:
    susie_results = pickle.load(f)

#create a list to store PIP data from all proteins
all_pip_data = []

#create a list to store credible set data
all_cs_data = []

#loop through each protein in the results
for protein_id, protein_data in susie_results.items():
    if 'pip' in protein_data:
        #get PIP dataframe and reset index to make variant_id a column
        pip_df = protein_data['pip'].reset_index()
        
        #add protein_id column
        pip_df['protein_id'] = protein_id
        
        #reorder columns to put protein_id first
        pip_df = pip_df[['protein_id', 'variant_id', 'pip', 'af']]
        
        #append to the list
        all_pip_data.append(pip_df)
    
    #extract credible sets
    if 'sets' in protein_data and 'cs' in protein_data['sets']:
        for cs_label, variant_indices in protein_data['sets']['cs'].items():
            #get the variants in this credible set
            if 'pip' in protein_data:
                variants = protein_data['pip'].index.values
                for idx in variant_indices:
                    if idx < len(variants):
                        cs_entry = {
                            'protein_id': protein_id,
                            'cs_label': cs_label,
                            'variant_id': variants[idx],
                        }
                        all_cs_data.append(cs_entry)

#combine all PIPs into one data frame
if all_pip_data:
    combined_pip_df = pd.concat(all_pip_data, ignore_index=True)
    
    #save file
    combined_pip_df.to_csv("~/ancestry_specific_folder/tensor_susie/0.05cis_pip_values.tsv", sep='\t', index=False)

#create credible sets dataframe
if all_cs_data:
    cs_df = pd.DataFrame(all_cs_data)
    
    #save file
    cs_df.to_csv("~/ancestry_specific_folder/tensor_susie/0.05credible_sets.tsv", sep='\t', index=False)
```

**Combine Cis Credible Sets and PIP Files**
```{python}
import pandas as pd

#read files
pip_df = pd.read_csv("~/ancestry_specific_folder/tensor_susie/0.05cis_pip_values.tsv", sep='\t')
cs_df = pd.read_csv("~/ancestry_specific_folder/tensor_susie/0.05cis_credible_sets.tsv", sep='\t')

#merge the credible sets with PIP values
merged_cs_df = cs_df.merge(
    pip_df[['protein_id', 'variant_id', 'pip', 'af']], 
    on=['protein_id', 'variant_id'], 
    how='left'
)

#save the merged data to a new file
merged_cs_df.to_csv("~/ancestry_specific_folder/tensor_susie/0.05cis_finemap.tsv", sep='\t', index=False)
```

**Create Trans Location File**
```{r}
library(dplyr)
library(data.table)
library(tidyr)

#read in file
data <- fread("~/ancestry_specific_folder/tensorqtl/trans/trans_fdr0.05.tsv ")

#ensure correct column names
colnames(data) <- c("variant_id","phenotype_id","pval","b","b_se","r2","af","beta_shape1","beta_shape2","true_df","pval_true_dof","pval_perm","pval_beta","qval","pval_nominal_threshold")

#isolate chr, pos, allele components
split <- data %>%
  separate(variant_id, into=c("chr", "pos", "A1", "A2"), sep=":")

#keep only needed columns
filtered <- split %>%
  select(phenotype_id, chr, pos)

#save files
write.table(filtered, "~/ancestry_specific_folder/0.05trans_locations.tsv", col.names=TRUE, row.names=FALSE, quote=FALSE, sep="\t")
```

**Trans Fine-Mapping**
```{bash}
#!/bin/bash

#paths
PLINK_PATH="~/ancestry_specific_folder/ancestry_final"
BED_PATH="~/ancestry_specific_folder/tensor_lm_table.bed.gz"
COV_PATH="~/ancestry_specific_folder/dummy_covariates.txt"
LOC_FILE="~/ancestry_specific_folder/0.05trans_locations.tsv"
OUT_DIR="~/ancestry_specific_folder/tensor_susie/0.05trans"
ERROR_FILE="~/ancestry_specific_folder/tensor_susie/trans_errors.tsv"

#create directories
mkdir -p "$OUT_DIR"
mkdir -p "$OUT_DIR/0.05phenotype_subsets"

#error log
> "$ERROR_FILE"
echo -e "phenotype_id\tchromosome\terror_message" > "$ERROR_FILE"

#extract header from locations file
HEADER=$(head -1 "$LOC_FILE")

#extract unique phenotypes
echo "Extracting unique phenotypes..."
unique_phenotypes=$(awk 'NR>1 {print $1}' "$LOC_FILE" | sort | uniq)

#process one phenotype at a time
for phenotype in $unique_phenotypes; do
    echo "Processing $phenotype..."
    
    #temp file with all locations (X to 23)
    TEMP_FILE=$(mktemp)
    awk -v phenotype="$phenotype" '
        NR>1 && $1==phenotype {
            # Convert X to chr23 in the chromosome column
            if ($2=="X" || $2=="chrX") {
                $2="chr23"
            }
            print $0
        }
    ' "$LOC_FILE" > "$TEMP_FILE"
    
    #unique chromosomes
    unique_chromosomes=$(awk '{print $2}' "$TEMP_FILE" | sort | uniq)
    
    #process each chromosome separately
    for chromosome in $unique_chromosomes; do
        echo "  Processing chromosome $chromosome..."
        
        #create phenotype-chromosome specific subset file
        CHR_SUBSET_FILE="$OUT_DIR/0.05phenotype_subsets/${phenotype}_${chromosome}_locations.tsv"
        
        #add header to the subset file
        echo -e "phenotype_id\tchr\tpos" > "$CHR_SUBSET_FILE"

        #add rows for this phenotype and chromosome, creating the COMBINED ID format
        awk -v phenotype="$phenotype" -v chromosome="$chromosome" '
            $2==chromosome {
                #output only phenotype_id, chr, and pos
                print $1 "\t" $2 "\t" $3
            }
        ' "$TEMP_FILE" >> "$CHR_SUBSET_FILE"
        
        #temporary output directory
        TMP_OUT_DIR=$(mktemp -d)
        
        #temporary file for error output
        ERROR_CAPTURE=$(mktemp)
        
        #run tensorqtl
        python3 -m tensorqtl \
            "$PLINK_PATH" \
            "$BED_PATH" \
            "$TMP_OUT_DIR/output" \
            --covariates "$COV_PATH" \
            --susie_loci "$CHR_SUBSET_FILE" \
            --mode trans_susie \
            --batch_size 10000 \
            --pval_threshold 1 \
            --maf_threshold 0.01 \
            --window 1000000 \
            --seed 123 > /dev/null 2> "$ERROR_CAPTURE"
        
        #check if there was an error
        if [ $? -ne 0 ]; then
            #get error message
            ERROR_MSG=$(cat "$ERROR_CAPTURE" | tr '\n' ' ' | tr '\t' ' ')
            
            #log the error to the error file
            echo -e "$phenotype\t$chromosome\t$ERROR_MSG" >> "$ERROR_FILE"
        else
            #copy only the result files to final directory
            mkdir -p "$OUT_DIR/${phenotype}_${chromosome}"
            
            #move any non-.log files
            find "$TMP_OUT_DIR" -type f -not -name "*.log" -exec cp {} "$OUT_DIR/${phenotype}_${chromosome}/" \;
        fi
        
        # Clean up temporary files and directories
        rm "$ERROR_CAPTURE"
        rm -rf "$TMP_OUT_DIR"
    done
    
    # Clean up temporary file
    rm "$TEMP_FILE"
    
    echo "Completed all chromosomes for $phenotype"
    echo "----------------------------------------"
done

echo "All phenotypes and chromosomes processed"
```

**Unpack Trans Susie Pickles**
```{python}
import pandas as pd
import pickle
import os
import glob
import re
import traceback

#root directory and pattern
root_dir = "~/ancestry_specific_folder/tensor_susie/0.05trans"
pickle_pattern = os.path.join(root_dir, "0.05_OID*_chr*/output.SuSiE.pickle")
all_pickle_files = glob.glob(pickle_pattern)

#error tracking
error_log = []
failed_phenotypes = []

#process each file
print(f"Processing {len(all_pickle_files)} files...")

for pickle_path in all_pickle_files:
    try:
        #extract OID and chromosome
        match = re.search(r'pq_OID(\d+)_chr(\d+)', pickle_path)
        if not match:
            error_log.append(f"Could not extract OID and chromosome from path: {pickle_path}")
            failed_phenotypes.append(os.path.basename(pickle_path))
            continue
            
        oid, chrom = match.group(1), match.group(2)

        #define output directory and filenames
        output_dir = os.path.dirname(pickle_path)
        pip_output_path = os.path.join(output_dir, f"OID{oid}_chr{chrom}_trans_pips.tsv")
        cs_output_path = os.path.join(output_dir, f"OID{oid}_chr{chrom}_trans_cs.tsv")
        finemap_output_path = os.path.join(output_dir, f"OID{oid}_chr{chrom}_finemap.tsv")
        
        #load the pickle file
        with open(pickle_path, 'rb') as f:
            trans_susie_results = pickle.load(f)

        #create lists to store PIP data and credible sets
        all_pip_data = []
        all_cs_data = []

        #process the TensorQTL trans-SuSiE output
        for protein_id, protein_data in trans_susie_results.items():
            if isinstance(protein_data, dict):
                #extract PIP values
                if 'pip' in protein_data:
                    pip_df = protein_data['pip']
                    
                    #handle different PIP data structures
                    if isinstance(pip_df, pd.DataFrame):
                        pip_df = pip_df.reset_index()
                    elif isinstance(pip_df, pd.Series):
                        pip_df = pip_df.reset_index(name='pip')
                    
                    #add protein_id column
                    pip_df['protein_id'] = protein_id
                    
                    #ensure we have the right columns
                    if 'variant_id' not in pip_df.columns and 'index' in pip_df.columns:
                        pip_df = pip_df.rename(columns={'index': 'variant_id'})
                    
                    #add allele frequency if available
                    if 'af' in pip_df.columns:
                        pip_df = pip_df[['protein_id', 'variant_id', 'pip', 'af']]
                    else:
                        pip_df = pip_df[['protein_id', 'variant_id', 'pip']]
                    
                    all_pip_data.append(pip_df)
                
                #extract credible sets
                if 'sets' in protein_data and 'cs' in protein_data['sets']:
                    for cs_label, variant_indices in protein_data['sets']['cs'].items():
                        if 'pip' in protein_data:
                            variants = None
                            pip_data = protein_data['pip']
                            
                            if isinstance(pip_data, (pd.DataFrame, pd.Series)):
                                variants = pip_data.index.values
                            
                            if variants is not None:
                                for idx in variant_indices:
                                    if idx < len(variants):
                                        cs_entry = {
                                            'protein_id': protein_id,
                                            'cs_label': cs_label,
                                            'variant_id': variants[idx],
                                        }
                                        all_cs_data.append(cs_entry)

        #create and save PIP dataframe
        if all_pip_data:
            combined_pip_df = pd.concat(all_pip_data, ignore_index=True)
            combined_pip_df.to_csv(pip_output_path, sep='\t', index=False)
        else:
            combined_pip_df = None

        #create and save credible sets dataframe
        if all_cs_data:
            cs_df = pd.DataFrame(all_cs_data)
            cs_df.to_csv(cs_output_path, sep='\t', index=False)
        else:
            cs_df = None

        #create the merged finemap file if both PIP and CS data exist
        if combined_pip_df is not None and cs_df is not None:
            pip_columns = ['protein_id', 'variant_id', 'pip']
            if 'af' in combined_pip_df.columns:
                pip_columns.append('af')
            
            merged_cs_df = cs_df.merge(
                combined_pip_df[pip_columns],
                on=['protein_id', 'variant_id'],
                how='left'
            )
            merged_cs_df.to_csv(finemap_output_path, sep='\t', index=False)
            print(finemap_output_path)
    
    except Exception as e:
        error_msg = f"Error processing OID{oid} Chr{chrom}: {str(e)}\n{traceback.format_exc()}"
        error_log.append(error_msg)
        failed_phenotypes.append(f"OID{oid}_chr{chrom}")

#write error log if there are any errors
if error_log:
    log_dir = "/home/claudia/susie/ancestry"
    os.makedirs(log_dir, exist_ok=True)
    error_log_file = os.path.join(log_dir, "susie_errors.log")
    
    with open(error_log_file, 'w') as f:
        f.write("SuSiE Processing Errors\n")
        f.write(f"Failed phenotypes: {len(failed_phenotypes)}\n\n")
        f.write("="*80 + "\n\n")
        
        for i, error_entry in enumerate(error_log):
            f.write(f"Error {i+1}:\n{error_entry}\n\n" + "-"*50 + "\n\n")
        
        f.write("Failed phenotypes:\n")
        for phenotype in failed_phenotypes:
            f.write(f"- {phenotype}\n")
    
    print(f"Finished with {len(failed_phenotypes)} errors. See log: {error_log_file}")
else:
    print("All files processed successfully.")
```

**Ensure Only Unique Credible Sets are Kept for Each File**
- If there are overlapping SNP/Protein pairs, keep the first one
Example:
protein_id	cs_label	variant_id	pip	af
OID20079_1	L1	chr4:16865080:G:A	0.5	0.012204724
OID20079_1	L1	chr4:16871683:G:A	0.5	0.012204724
OID20079_2	L1	chr4:16865080:G:A	0.5	0.012204724
OID20079_2	L1	chr4:16871683:G:A	0.5	0.012204724
to:
protein_id	cs_label	variant_id	pip	af
OID20079_1	L1	chr4:16865080:G:A	0.5	0.012204724
OID20079_1	L1	chr4:16871683:G:A	0.5	0.012204724

```{r}
library(dplyr)
library(tidyr)
library(stringr)

#root directory
root_dir <- "/home/claudia/tensor_susie/all_race/0.05trans"

#find all finemap.tsv files
finemap_files <- list.files(path = root_dir, pattern = "*finemap.tsv$", recursive = TRUE, full.names = TRUE)
cat("Found", length(finemap_files), "finemap.tsv files\n")

#create an error log data frame
error_log <- data.frame(
  file_path = character(),
  error_message = character(),
  timestamp = character(),
  stringsAsFactors = FALSE
)

#process each file
for (file_path in finemap_files) {
  #extract file info for reporting
  file_name <- basename(file_path)
  dir_name <- basename(dirname(file_path))
  output_path <- file.path(dirname(file_path), paste0(tools::file_path_sans_ext(file_name), "_condensed.tsv"))
  
  cat("\nProcessing:", dir_name, "/", file_name, "\n")
  
  #read the file
  tryCatch({
    finemap_data <- read.table(file_path, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
    
    if (nrow(finemap_data) == 0) {
      cat("  Empty file, skipping\n")
      #add to error log
      error_log <- rbind(error_log, data.frame(
        file_path = file_path,
        error_message = "Empty file",
        timestamp = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
        stringsAsFactors = FALSE
      ))
      next
    }
      
    #keep the first occurrence of each unique variant pattern
    first_occurrences <- finemap_data %>%
      group_by(variant_id) %>%
      slice(1) %>%
      ungroup() %>%
      arrange(as.numeric(str_extract(gene_id, "\\d+$"))) %>%
      select(gene_id, cs_label, variant_id, pip, af)

    
    #write the condensed data to file
    write.table(first_occurrences, file = output_path, sep = "\t", 
                row.names = FALSE, quote = FALSE)
    cat("Condensed file saved to:", output_path, "\n")
    
  }, error = function(e) {
    error_message <- e$message
    cat("ERROR processing file:", error_message, "\n")
    
    #add to error log
    error_log <- rbind(error_log, data.frame(
      file_path = file_path,
      error_message = error_message,
      timestamp = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
      stringsAsFactors = FALSE
    ))
  })
}

#write all errors to a single file
error_log_path <- file.path(root_dir, "processing_errors.tsv")
write.table(error_log, file = error_log_path, sep = "\t", row.names = FALSE, quote = FALSE)
cat("\nError log saved to:", error_log_path, "\n")
cat("Total errors:", nrow(error_log), "\n")
```

**Combine Individual Protein Fine-Mapped Files into One**
```{r}
library(dplyr)
library(stringr)

#root directory where condensed finemap files are located
root_dir <- "~/ancestry_specific_folder/tensor_susie/0.05trans"

#find all condensed finemap files
condensed_files <- list.files(path = root_dir, pattern = "*_condensed.tsv$", recursive = TRUE, full.names = TRUE)
cat("Found", length(condensed_files), "condensed finemap files\n")

#process all files and combine using rbind
all_data <- list()
for (i in 1:length(condensed_files)) {
  file <- condensed_files[i]
  tryCatch({
    #read the file
    data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
    
    #keep only the columns we want
    if (all(c("protein_id", "cs_label", "variant_id", "pip", "af") %in% colnames(data))) {
      data <- data[, c("protein_id", "cs_label", "variant_id", "pip", "af")]
      all_data[[i]] <- data
      cat("Added data from:", file, "\n")
    } else {
      cat("File missing required columns, skipping:", file, "\n")
    }
  }, error = function(e) {
    cat("ERROR processing file:", file, ":", e$message, "\n")
  })
}

#combine all successful results
combined_results <- do.call(rbind, all_data)

#count unique protein_id and cs_label pairs
protein_cs_pairs <- combined_results %>%
  select(protein_id, cs_label) %>%
  distinct()

num_unique_pairs <- nrow(protein_cs_pairs)

#write the combined results to a file
combined_file_path <- file.path("~/ancestry_specific_folder/tensor_susie/", "0.05trans_finemap.tsv")
write.table(combined_results, file = combined_file_path, sep = "\t", row.names = FALSE, quote = FALSE)
```

**Filter Final File**
- Group by OID_#, credible set label, and chromosome to isolate credible sets
- Keep only credible sets with greater than or equal to 0.9 coverage
- Relabel credible set labels to make each unique within OID group
- Remove _# after OID#####
```{r}
library(dplyr)
library(data.table)
library(stringr)

trans_condensed <- fread("~/ancestry_specific_folder/tensor_susie/0.05trans_finemap.tsv", header=TRUE)
colnames(trans_condensed) <- c("gene_id", "cs_label", "variant_id", "pip", "af")

#add chr column
trans_condensed$chr <- str_extract(trans_condensed$variant_id, "chr\\d+")
trans_condensed$chr[is.na(trans_condensed$chr)] <- "chr23"
trans_condensed$base_gene_id <- str_extract(trans_condensed$gene_id, "^[^_]+")

#keep only credible sets that add to .9 (conservative filter)
trans_filtered <- trans_condensed %>%
  group_by(gene_id, cs_label, chr) %>%
  filter(sum(pip) >= 0.9) %>%
  ungroup() %>%
  mutate(
    #create a unique identifier for each credible set
    cs_combo = paste(gene_id, cs_label, chr, sep = "_"),
    #assign ID based on first appearance order
    cs_id = as.numeric(factor(cs_combo, levels = unique(cs_combo)))
  ) %>%
  select(-cs_combo)

#relabel credible sets to make unique cs_label per protein
trans_cs_relabel <- trans_filtered %>%
  mutate(original_order = row_number()) %>%
  group_by(base_gene_id) %>%
  mutate(
    #assign sequential L numbers based on unique cs_id values
    new_cs_label = paste0("L", dense_rank(cs_id))
  ) %>%
  ungroup() %>%
  arrange(original_order) %>%
  select(-cs_label, -original_order) %>%
  rename(cs_label = new_cs_label)

#rearrange
trans_reorder <- trans_cs_relabel %>%
  select(base_gene_id, cs_label, variant_id, pip)
colnames(trans_reorder) <- c("gene_id", "cs_label", "variant_id", "pip")

#save
write.table(trans_reorder, "~/ancestry_specific_folder/tensor_susie/0.05trans_finemap_edited.tsv", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
```

