---
title: "Validate MESA Credible Sets in UKB"
author: "Claudia Krueger"
date: "2026-05-23"
output: html_document
---

**Extract Cis Regions from UK Biobank pQTL Summary Statistics (hg37) Based on MESA pQTL Window Assignments (hg38)**
```{bash}
#!/usr/bin/env bash
set -euo pipefail

LOC="~/ukb_finemap/1kg/allchr_protein_location_file_ancestry"
IN_DIR="~/ukbio/ancestry_specific_folder"
OUT_DIR="~/ukbio/ancestry_cis_sumstats"
FLANK=1000000

mkdir -p "$OUT_DIR"

tail -n +2 "$LOC" | while read -r OID CHR LEFT RIGHT; do
  LEFTW=$((LEFT-FLANK))
  RIGHTW=$((RIGHT+FLANK))

  FILE=$(ls "$IN_DIR"/*_"$OID"_* 2>/dev/null | head -n1 || true)

  if [[ -z "$FILE" ]]; then
    echo "$OID → no file"
    continue
  fi

  OUT="$OUT_DIR/$OID.tsv"

  #filter by GENPOS ($2). Add POS_hg37 from ID ($3).
  zcat "$FILE" | awk -v FS='\t' -v OFS='\t' -v chr="$CHR" -v lo="$LEFTW" -v hi="$RIGHTW" '
    NR==1 {
      print $0, "POS_hg37"
      next
    }
    {
      split($3, a, ":")
      pos_hg37 = (length(a[2]) ? a[2] : "NA")
      # filter uses GENPOS column ($2)
      if ($1==chr && $2>=lo && $2<=hi) print $0, pos_hg37
    }
  ' > "$OUT"

  if [[ $(wc -l < "$OUT") -le 1 ]]; then
    rm -f "$OUT"
    echo "$OID → no variants in window"
  else
    echo "$OID → success"
  fi
done
```

**Download 1KG reference genome files (hg37)**
```{bash}
base="https://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502"

cd ~/ukb_finemap/1kg

#autosomes
for c in {1..22}; do
  wget -c "$base/ALL.chr${c}.phase3_shapeit2_mvncall_integrated_v5b.20130502.genotypes.vcf.gz" \
  
done

#sex chr + MT
wget -c "$base/ALL.chrX.phase3_shapeit2_mvncall_integrated_v1c.20130502.genotypes.vcf.gz" 

#pannel
wget "https://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20130502.ALL.panel"
```

**Filter 1KG Reference Files to SNPs Present in UKB for Each Ancestry**
```{bash}
#!/usr/bin/env bash
set -euo pipefail

IN="~/1000g_phase3_20130502"
OUT="~/ukb_finemap/1kg"         
mkdir -p "$OUT"

PANEL="$OUT/integrated_call_samples_v3.20130502.ALL.panel"

#phase 3 filenames (20130502 release)
AUTO="ALL.chr%d.phase3_shapeit2_mvncall_integrated_v5b.20130502.genotypes.vcf.gz"
X="ALL.chrX.phase3_shapeit2_mvncall_integrated_v1c.20130502.genotypes.vcf.gz"

for SP in AFR AMR EAS EUR; do
  echo "== $SP =="

  SAMPLES="$OUT/$SP.samples.txt"
  awk -F'\t' -v sp="$SP" 'NR>1 && $3==sp {print $1}' "$PANEL" > "$SAMPLES"

  TMP="$OUT/tmp_$SP"
  mkdir -p "$TMP"

  #subset chr1-22
  for c in $(seq 1 22); do
    bcftools view -S "$SAMPLES" -Oz -o "$TMP/$SP.chr$c.vcf.gz" "$IN/$(printf "$AUTO" "$c")"
    tabix -p vcf "$TMP/$SP.chr$c.vcf.gz"
  done

  #subset chrX
  bcftools view -S "$SAMPLES" -Oz -o "$TMP/$SP.chrX.vcf.gz" "$IN/$X"
  tabix -p vcf "$TMP/$SP.chrX.vcf.gz"

  #concat to one file
  bcftools concat -Oz -o "$OUT/$SP.ALL_CHR.vcf.gz" \
    "$TMP/$SP.chr"{1..22}".vcf.gz" "$TMP/$SP.chrX.vcf.gz"
  tabix -p vcf "$OUT/$SP.ALL_CHR.vcf.gz"

  echo "wrote $OUT/$SP.ALL_CHR.vcf.gz"
done
```

**Build PLINK and .bgen files**
```{bash}
#build ID COLUMN CHR:POS:REF:ALT
plink2 --pfile ~/ukb_finemap/1kg/EUR --set-all-var-ids '@:#:$r:$a' --snps-only --make-pgen --out ~/ukb_finemap/1kg/EUR_id
plink2 --pfile ~/ukb_finemap/1kg/AFR --set-all-var-ids '@:#:$r:$a' --snps-only --make-pgen --out ~/ukb_finemap/1kg/AFR_id
plink2 --pfile ~/ukb_finemap/1kg/EAS --set-all-var-ids '@:#:$r:$a' --snps-only --make-pgen --out ~/ukb_finemap/1kg/EAS_id
plink2 --pfile ~/ukb_finemap/1kg/AMR --set-all-var-ids '@:#:$r:$a' --snps-only --make-pgen --out ~/ukb_finemap/1kg/AMR_id

#final single ancestry files
plink2 --pfile ~/ukb_finemap/1kg/EUR_id --max-alleles 2 --make-pgen --out ~/ukb_finemap/1kg/EUR_final
plink2 --pfile ~/ukb_finemap/1kg/AFR_id --max-alleles 2 --make-pgen --out ~/ukb_finemap/1kg/AFR_final
plink2 --pfile ~/ukb_finemap/1kg/EAS_id --max-alleles 2 --make-pgen --out ~/ukb_finemap/1kg/EAS_final
plink2 --pfile ~/ukb_finemap/1kg/AMR_id --max-alleles 2 --make-pgen --out ~/ukb_finemap/1kg/AMR_final

#build .bgen files
plink2 --pfile ~/ukb_finemap/1kg/EUR_final --export bgen-1.2 --out ~/ukb_finemap/1kg/EUR_final
plink2 --pfile ~/ukb_finemap/1kg/AFR_final --export bgen-1.2 --out ~/ukb_finemap/1kg/AFR_final
plink2 --pfile ~/ukb_finemap/1kg/EAS_final --export bgen-1.2 --out ~/ukb_finemap/1kg/EAS_final
plink2 --pfile ~/ukb_finemap/1kg/AMR_final --export bgen-1.2 --out ~/ukb_finemap/1kg/AMR_final

#build .bgen index files
~/multi_susie/bgen/build/apps/bgenix -g ~/ukb_finemap/1kg/EUR_final.bgen -index
~/multi_susie/bgen/build/apps/bgenix -g ~/ukb_finemap/1kg/AFR_final.bgen -index
~/multi_susie/bgen/build/apps/bgenix -g ~/ukb_finemap/1kg/EAS_final.bgen -index
~/multi_susie/bgen/build/apps/bgenix -g ~/ukb_finemap/1kg/AMR_final.bgen -index
```

**Build Z Files from 1KG**
```{bash}
#!/usr/bin/env bash
set -euo pipefail

#inputs
PROT_DIR="~/ukbio/ancestry_cis_sumstats"
PVAR="~/ukb_finemap/1kg/ancestry_final.pvar"   # no chr prefix; X coded as X

#output
OUT_DIR="~/ukb_finemap/1kg/z_files/EUR"
mkdir -p "$OUT_DIR"

#make file list
FILE_LIST="~/ukb_finemap/1kg/ancestry_protein_files.txt"
ls "$PROT_DIR"/*.tsv.gz > "$FILE_LIST"
echo "Total protein files:"
wc -l "$FILE_LIST"
echo
i=0
while read -r f; do
  ((++i))
  base="$(basename "$f")"
  stem="${base%.tsv.gz}"
  out_z="$OUT_DIR/${stem}.z"
  echo "[$i] $stem"
  #build z by:
  # 1) reading protein sumstats IDs (col 3) and NORMALIZING them:
  #   chr1  -> 1
  #   chrX  -> X
  #   chr23 -> 23
  # 2) scanning pvar once for matches and printing pvar-derived fields
  awk -v FS='\t' -v OFS=' ' '
    BEGIN { print "rsid chromosome position allele1 allele2" }
    #first input = protein file stream from zcat
    NR==FNR {
      # skip header if present
      if (FNR==1 && $1=="CHROM") next
      id=$3
      sub(/^chr/, "", id)
      sub(/^23:/, "X:", id)
      keep[id]=1
      next
    }
    #second input = pvar
    /^##/ { next }
    /^#CHROM/ { next }
    ($3 in keep) {
      # pvar columns: CHROM POS ID REF ALT
      print $3, $1, $2, $5, $4
    }
  ' <(zcat "$f") "$PVAR" > "$out_z"
done < "$FILE_LIST"
echo
echo "done"
```

**Ensure no Duplicate SNPs in Z Files**
```{r}
library(data.table)
library(dplyr)

dup_proteins <- c(
  "OID20203",
  "OID20239",
  "OID20297",
  "OID20360",
  "OID20403",
  "OID20442",
  "OID20547",
  "OID20573",
  "OID20579",
  "OID20609",
  "OID20710",
  "OID21228",
  "OID21281",
  "OID21306",
  "OID21436",
  "OID30208",
  "OID30223",
  "OID30672",
  "OID30689",
  "OID30705",
  "OID30730",
  "OID30757",
  "OID30779",
  "OID30927",
  "OID31087",
  "OID31274",
  "OID31362",
  "OID31365"
)

for (protein in dup_proteins){
  file_name <- paste0("~/ukb_finemap/1kg/z_files/ancestry_specific_folder/", protein, ".z")
  file <- fread(file_name)
  file_unique <- file %>% unique()
  write.table(file_unique, file_name, sep=" ", row.names=FALSE, quote=FALSE)
}
```

**Build Master File**
```{bash}
#!/bin/bash

Z_DIR="~/ukb_finemap/1kg/z_files/ancestry_specific_folder"
MASTER="~/ukb_finemap/1kg/master_files/ancestry_master.txt"

#create master file with header
echo "z;bgen;bgi;bcor;n_samples" > "$MASTER"

count=0

for Z_FILE in "$Z_DIR"/*.z; do

    # skip if no files found
    [[ -e "$Z_FILE" ]] || continue

    olinkid=$(basename "$Z_FILE" .z)

    echo "${Z_FILE};~/ukb_finemap/1kg/ancestry_final.bgen;~/ukb_finemap/1kg/ancestry_final.bgen.bgi;~/ukb_finemap/1kg/ld_files/ancestry_specific_folder/${olinkid}.bcor;661" >> "$MASTER"

    ((count++))

done
```

**Build LD Files Using LDStore**
```{bash}
nohup ~/multi_susie/ldstore_v2.0_x86_64/ldstore_v2.0_x86_64 --in-files ~/ukb_finemap/1kg/master_files/ancestry_master.txt --write-bcor --read-only-bgen --n-threads 1 > ~/ukb_finemap/ancestry_ldstore.log 2>&1 & 
```

**Wrapper for Fine-mapping Runs**
```{bash}
#!/bin/bash

source "$(conda info --base)/etc/profile.d/conda.sh

OlinkID=$1
chr=$2
left=$3
right=$4
LOG="${5:-~/ukb_finemap/log_files/runtime_log.txt}"
SUSHIE_DONE="${6:-}"

mkdir -p "$(dirname "$LOG")"
if [[ ! -f "$LOG" ]]; then
    echo -e "OlinkID\tmethod\telapsed_sec" > "$LOG"
fi

echo "Processing: ${OlinkID}"

#check .bcor file exists for each pop before building LD
missing_bcor=0
for pop in EUR AFR AMR EAS; do
    if [[ ! -f "~/ukb_finemap/1kg/ld_files/${pop}/${OlinkID}.bcor" ]]; then
        echo "  Missing .bcor file for ${pop} - skipping ${OlinkID}"
        missing_bcor=1
        break
    fi
done
[[ $missing_bcor -eq 1 ]] && exit 0

#make LD file for each population
echo "  Building ${OlinkID} Individual LD Files..."
python ~/ukb_finemap/map_scripts/01build_individual_ld.py --oid "$OlinkID"

#build meta LD file
echo "  Building ${OlinkID} META LD File..."
Rscript ~/ukb_finemap/map_scripts/02build_meta_ld.R --oid "$OlinkID"

#filter and sort summary stats files
echo "  Building Summary Statistic Files..."
Rscript ~/ukb_finemap/map_scripts/03filter_sumstats.R --oid "$OlinkID"

#run SuSiER
echo "  Running ${OlinkID} SuSiEr..."
t0=$(date +%s)
Rscript ~/ukb_finemap/map_scripts/04susier.R --oid "$OlinkID" 2>&1 \
    | grep -v "Running SuSiEr" \
    | grep -v "__HINT__" \
    | grep -v "estimate_residual_variance" \
    | grep -v "Also note, when covariates"

exit_status=$?
if [ $exit_status -eq 0 ]; then
    duration=$(( $(date +%s) - t0 ))
else
    duration="NA"
fi
echo -e "${OlinkID}\tsusier\t${duration}" >> "$LOG"

#run SuShiE
echo "  Running ${OlinkID} SuShiE..."
t0=$(date +%s)
conda run -n sushie310 sushie finemap \
    --summary \
    --ld-adjust 0.02 \
    --platform gpu \
    --gwas \
        ~/ukb_finemap/tmp_sumstats/EUR/${OlinkID}.txt \
        ~/ukb_finemap/tmp_sumstats/AFR/${OlinkID}.txt \
        ~/ukb_finemap/tmp_sumstats/AMR/${OlinkID}.txt \
        ~/ukb_finemap/tmp_sumstats/EAS/${OlinkID}.txt \
    --ld \
        ~/ukb_finemap/1kg/tmp_ld_files/EUR/${OlinkID}_common.txt \
        ~/ukb_finemap/1kg/tmp_ld_files/AFR/${OlinkID}_common.txt \
        ~/ukb_finemap/1kg/tmp_ld_files/AMR/${OlinkID}_common.txt \
        ~/ukb_finemap/1kg/tmp_ld_files/EAS/${OlinkID}_common.txt \
    --sample-size 33187 931 97 262 \
    --gwas-header CHROM ID POS_hg37 ALLELE1 ALLELE0 Z BETA SE Pvalue GENPOS A1FREQ \
    --trait "$OlinkID" \
    --output ~/ukb_finemap/sushie_output/${OlinkID} \
    --min-snps=40
    2>&1 | grep -v "FutureWarning" | grep -v "suffixes.*duplicate" | grep -v "snps_gwas = snps_gwas.merge"

exit_status=$?
if [ $exit_status -eq 0 ]; then
    duration=$(( $(date +%s) - t0 ))
else
    duration="NA"
fi
echo -e "${OlinkID}\tsushie\t${duration}" >> "$LOG"

#signal to launcher that SuShiE is done so the next job can start
[[ -n "$SUSHIE_DONE" ]] && touch "$SUSHIE_DONE"

#run MultiSuSiE
echo "  Running ${OlinkID} MultiSuSiE..."
t0=$(date +%s)
conda run -n MultiSuSiE python ~/ukb_finemap/map_scripts/05multi_susie.py --oid "$OlinkID"

exit_status=$?
if [ $exit_status -eq 0 ]; then
    duration=$(( $(date +%s) - t0 ))
else
    duration="NA"
fi
echo -e "${OlinkID}\tmulti_susie\t${duration}" >> "$LOG"

#run SuSiEx
wait $!
echo "  Running ${OlinkID} SuSiEx..."
python ~/ukb_finemap/map_scripts/06susiex_ld_calc.py --oid "$OlinkID"

t0=$(date +%s)
~/SuSiEx/bin/SuSiEx \
    --sst_file=~/ukb_finemap/tmp_sumstats/EUR/${OlinkID}.txt,~/ukb_finemap/tmp_sumstats/AFR/${OlinkID}.txt,~/ukb_finemap/tmp_sumstats/AMR/${OlinkID}.txt,~/ukb_finemap/tmp_sumstats/EAS/${OlinkID}.txt \
    --n_gwas=33187,932,97,262 \
    --ld_file=~/ukb_finemap/1kg/tmp_ld_files/EUR/${OlinkID}_susiex,~/ukb_finemap/1kg/tmp_ld_files/AFR/${OlinkID}_susiex,~/ukb_finemap/1kg/tmp_ld_files/AMR/${OlinkID}_susiex,~/ukb_finemap/1kg/tmp_ld_files/EAS/${OlinkID}_susiex \
    --out_dir=~/ukb_finemap/susiex_output \
    --out_name="${OlinkID}" \
    --chr=${chr} \
    --bp=0,100000000000000000000 \
    --chr_col=1,1,1,1 \
    --snp_col=2,2,2,2 \
    --bp_col=3,3,3,3 \
    --a1_col=4,4,4,4 \
    --a2_col=5,5,5,5 \
    --eff_col=7,7,7,7 \
    --se_col=8,8,8,8 \
    --pval_col=9,9,9,9 \
    --maf=0 \
    --n_sig=10 \
    --keep-ambig=TRUE \
    --mult-step=TRUE \
    --tol=0.001 \
    --plink=~/SuSiEx/utilities/plink

exit_status=$?
if [ $exit_status -eq 0 ]; then
    duration=$(( $(date +%s) - t0 ))
else
    duration="NA"
fi
echo -e "${OlinkID}\tsusiex\t${duration}" >> "$LOG"

#clean up ld files
rm ~/ukb_finemap/1kg/tmp_ld_files/*/${OlinkID}_full.txt
rm ~/ukb_finemap/1kg/tmp_ld_files/*/${OlinkID}_common.txt
rm ~/ukb_finemap/1kg/tmp_ld_files/common_snp_lists/${OlinkID}.txt
rm ~/ukb_finemap/tmp_sumstats/*/${OlinkID}.txt

echo "Done: ${OlinkID}"
```

**01build_individual_ld.py**
```{python}
import numpy as np
import pandas as pd
import sys
import os
import argparse

#argparse
def check_arg(args=None):
    parser = argparse.ArgumentParser(description="Build LD files for all proteins")
    parser.add_argument('--oid', help='OlinkID', required=True)
    return parser.parse_args(args)

args = check_arg(sys.argv[1:])
protein_id = args.oid

ancestries = ['EUR', 'AFR', 'EAS', 'AMR']

#build and save ld matricies
def build_ld(protein_id, ldstore_path="~/multi_susie/ldstore_v2.0_x86_64/ldstore_v2.0_x86_64"):

    work_dir = f'~/ukb_finemap/tmp/{protein_id}'
    os.makedirs(work_dir, exist_ok=True)
    original_dir = os.getcwd()
    os.chdir(work_dir)

    for ancestry in ancestries:
        bcor_file = f'~/ukb_finemap/1kg/ld_files/{ancestry}/{protein_id}.bcor'
        tmp_ld_file = f'{work_dir}/{protein_id}_{ancestry}.ld'

        #output path — what R reads
        output_dir = f'~/ukb_finemap/1kg/tmp_ld_files/{ancestry}'
        os.makedirs(output_dir, exist_ok=True)
        output_file = f'{output_dir}/{protein_id}_full.txt'

        try:
            cmd = f"{ldstore_path} --bcor-to-text --bcor-file {bcor_file} --ld-file {tmp_ld_file} > /dev/null 2>&1"

            #retry ldstore up to 3 times with LD matrix validation
            max_attempts = 3
            ld_matrix = None
            for attempt in range(max_attempts):
                os.system(cmd)
                if not os.path.exists(tmp_ld_file) or os.path.getsize(tmp_ld_file) == 0:
                    print(f"Retry {attempt + 1}/{max_attempts} — empty file for {protein_id} {ancestry}")
                    continue
                try:
                    ld_matrix = np.loadtxt(tmp_ld_file)
                    if ld_matrix.ndim != 2 or ld_matrix.shape[0] != ld_matrix.shape[1]:
                        raise ValueError(f"LD matrix not square: {ld_matrix.shape}")
                    break
                except Exception as e:
                    print(f"Retry {attempt + 1}/{max_attempts} — malformed file for {protein_id} {ancestry}: {e}")
                    ld_matrix = None
                    if os.path.exists(tmp_ld_file):
                        os.remove(tmp_ld_file)

            if ld_matrix is None:
                print(f"Error: ldstore failed after {max_attempts} attempts for {protein_id} {ancestry}")
                os.chdir(original_dir)
                return False

            #save as tab-separated text file for R to read
            pd.DataFrame(ld_matrix).to_csv(output_file, sep='\t', index=False, header=False)
            print(f"Saved LD matrix for {protein_id} {ancestry} -> {output_file}")

        except Exception as e:
            print(f"Error processing LD for {protein_id} {ancestry}: {e}")
            os.chdir(original_dir)
            return False
        finally:
            if os.path.exists(tmp_ld_file):
                os.remove(tmp_ld_file)

    os.chdir(original_dir)
    try:
        os.rmdir(work_dir)
    except:
        pass

    return True

#main
success = build_ld(protein_id)
if not success:
    print(f"FAILED: Could not build LD files for {protein_id}")
    sys.exit(1)

print(f"SUCCESS: Built LD files for {protein_id}")
sys.exit(0)
```

**02build_meta_ld.py**
```{python}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(argparse))

#setup command line argument parsing
parser <- ArgumentParser()
parser$add_argument("--oid", required=TRUE, help="OID number (e.g., OID20089)")
args <- parser$parse_args()

protein_id <- args$oid

eur_snp_file <- fread(paste0("~/ukb_finemap/1kg/z_files/EUR/", protein_id, ".z"))
eur_matrix <- as.matrix(fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/EUR/", protein_id, "_full.txt")))
afr_snp_file <- fread(paste0("~/ukb_finemap/1kg/z_files/AFR/", protein_id, ".z"))
afr_matrix <- as.matrix(fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/AFR/", protein_id, "_full.txt")))
amr_snp_file <- fread(paste0("~/ukb_finemap/1kg/z_files/AMR/", protein_id, ".z"))
amr_matrix <- as.matrix(fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/AMR/", protein_id, "_full.txt")))
eas_snp_file <- fread(paste0("~/ukb_finemap/1kg/z_files/EAS/", protein_id, ".z"))
eas_matrix <- as.matrix(fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/EAS/", protein_id, "_full.txt")))

eur_snps <- eur_snp_file$rsid
afr_snps <- afr_snp_file$rsid
amr_snps <- amr_snp_file$rsid
eas_snps <- eas_snp_file$rsid

colnames(eur_matrix) <- eur_snps
row.names(eur_matrix) <- eur_snps
colnames(afr_matrix) <- afr_snps
row.names(afr_matrix) <- afr_snps
colnames(amr_matrix) <- amr_snps
row.names(amr_matrix) <- amr_snps
colnames(eas_matrix) <- eas_snps
row.names(eas_matrix) <- eas_snps

#identify SNPs with mostly NA in EUR LD
eur_na_prop <- rowSums(is.na(eur_matrix)) / ncol(eur_matrix)
eur_major_na_snps <- which(eur_na_prop > 0.9)
eur_major_na_names <- rownames(eur_matrix)[eur_major_na_snps]
afr_na_prop <- rowSums(is.na(afr_matrix)) / ncol(afr_matrix)
afr_major_na_snps <- which(afr_na_prop > 0.9)
afr_major_na_names <- rownames(afr_matrix)[afr_major_na_snps]
amr_na_prop <- rowSums(is.na(amr_matrix)) / ncol(amr_matrix)
amr_major_na_snps <- which(amr_na_prop > 0.9)
amr_major_na_names <- rownames(amr_matrix)[amr_major_na_snps]
eas_na_prop <- rowSums(is.na(eas_matrix)) / ncol(eas_matrix)
eas_major_na_snps <- which(eas_na_prop > 0.9)
eas_major_na_names <- rownames(eas_matrix)[eas_major_na_snps]

bad_snps_list <- list(eur_major_na_names, afr_major_na_names, amr_major_na_names, eas_major_na_names)

common_snps <- intersect(eur_snps, intersect(afr_snps, intersect(amr_snps, eas_snps)))
common_snps_clean <- Reduce(setdiff, c(list(common_snps), bad_snps_list))
common_snps_ordered <- intersect(colnames(eur_matrix), common_snps_clean)

eur_filtered <- eur_matrix[common_snps_ordered, common_snps_ordered]
afr_filtered <- afr_matrix[common_snps_ordered, common_snps_ordered]
amr_filtered <- amr_matrix[common_snps_ordered, common_snps_ordered]
eas_filtered <- eas_matrix[common_snps_ordered, common_snps_ordered]

common_snp_file <- paste0("~/ukb_finemap/1kg/tmp_ld_files/common_snp_lists/", protein_id, ".txt")
write.table(common_snps_ordered, common_snp_file, sep= "\n", row.names=FALSE, col.names=FALSE, quote=FALSE)

eur_matrix_file <- paste0("~/ukb_finemap/1kg/tmp_ld_files/EUR/", protein_id, "_common.txt")
write.table(eur_filtered, eur_matrix_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
afr_matrix_file <- paste0("~/ukb_finemap/1kg/tmp_ld_files/AFR/", protein_id, "_common.txt")
write.table(afr_filtered, afr_matrix_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
amr_matrix_file <- paste0("~/ukb_finemap/1kg/tmp_ld_files/AMR/", protein_id, "_common.txt")
write.table(amr_filtered, amr_matrix_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
eas_matrix_file <- paste0("~/ukb_finemap/1kg/tmp_ld_files/EAS/", protein_id, "_common.txt")
write.table(eas_filtered, eas_matrix_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)

eur_ancestry_weight <- (33187/(33187+931+97+262))
afr_ancestry_weight <- (932/(33187+931+97+262))
amr_ancestry_weight <- (97/(33187+931+97+262))
eas_ancestry_weight <- (262/(33187+931+97+262))

meta_matrix <- (
  eur_ancestry_weight * eur_filtered +
    afr_ancestry_weight * afr_filtered +
    amr_ancestry_weight * amr_filtered +
    eas_ancestry_weight * eas_filtered
)

meta_matrix_file <- paste0("~/ukb_finemap/1kg/tmp_ld_files/META/", protein_id, "_common.txt")
write.table(meta_matrix, meta_matrix_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
```

**03filter_sumstats.R**
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(argparse))

#setup command line argument parsing
parser <- ArgumentParser()
parser$add_argument("--oid", required=TRUE, help="OID number (e.g., OID20089)")
args <- parser$parse_args()

protein_id <- args$oid
#protein_id <- "OID20049"

meta_sumstats <- fread(paste0("~/ukbio/all_race_cis_sumstats/", protein_id, ".tsv.gz"))
eur_sumstats <- fread(paste0("~/ukbio/euro_amer_cis_sumstats/", protein_id, ".tsv.gz"))
afr_sumstats <- fread(paste0("~/ukbio/afr_amer_cis_sumstats/", protein_id, ".tsv.gz"))
amr_sumstats <- fread(paste0("~/ukbio/hisp_amer_cis_sumstats/", protein_id, ".tsv.gz"))
eas_sumstats <- fread(paste0("~/ukbio/chn_amer_cis_sumstats/", protein_id, ".tsv.gz"))

common_snps <- fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/common_snp_lists/", protein_id, ".txt"), header=FALSE)

meta_sumstats_filtered <- meta_sumstats %>%
  mutate(ID = gsub("^chr", "", ID)) %>%
  mutate(CHROM = gsub("^chr", "", CHROM)) %>%
  mutate(Z = BETA/SE) %>%
  filter(ID %in% common_snps$V1) %>%
  select(CHROM, ID, POS_hg37, ALLELE1, ALLELE0, Z, BETA, SE, Pvalue, GENPOS, A1FREQ)
eur_sumstats_filtered <- eur_sumstats %>%
  mutate(ID = gsub("^chr", "", ID)) %>%
  mutate(CHROM = gsub("^chr", "", CHROM)) %>%
  mutate(Z = BETA/SE) %>%
  filter(ID %in% common_snps$V1) %>%
  select(CHROM, ID, POS_hg37, ALLELE1, ALLELE0, Z, BETA, SE, Pvalue, GENPOS, A1FREQ)
afr_sumstats_filtered <- afr_sumstats %>%
  mutate(ID = gsub("^chr", "", ID)) %>%
  mutate(CHROM = gsub("^chr", "", CHROM)) %>%
  mutate(Z = BETA/SE) %>%
  filter(ID %in% common_snps$V1) %>%
  select(CHROM, ID, POS_hg37, ALLELE1, ALLELE0, Z, BETA, SE, Pvalue, GENPOS, A1FREQ)
amr_sumstats_filtered <- amr_sumstats %>%
  mutate(ID = gsub("^chr", "", ID)) %>%
  mutate(CHROM = gsub("^chr", "", CHROM)) %>%
  mutate(Z = BETA/SE) %>%
  filter(ID %in% common_snps$V1) %>%
  select(CHROM, ID, POS_hg37, ALLELE1, ALLELE0, Z, BETA, SE, Pvalue, GENPOS, A1FREQ)
eas_sumstats_filtered <- eas_sumstats %>%
  mutate(ID = gsub("^chr", "", ID)) %>%
  mutate(CHROM = gsub("^chr", "", CHROM)) %>%
  mutate(Z = BETA/SE) %>%
  filter(ID %in% common_snps$V1) %>%
  select(CHROM, ID, POS_hg37, ALLELE1, ALLELE0, Z, BETA, SE, Pvalue, GENPOS, A1FREQ)

meta_sumstats_file <- paste0("~/ukb_finemap/tmp_sumstats/META/", protein_id, ".txt")
write.table(meta_sumstats_filtered, meta_sumstats_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
eur_sumstats_file <- paste0("~/ukb_finemap/tmp_sumstats/EUR/", protein_id, ".txt")
write.table(eur_sumstats_filtered, eur_sumstats_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
afr_sumstats_file <- paste0("~/ukb_finemap/tmp_sumstats/AFR/", protein_id, ".txt")
write.table(afr_sumstats_filtered, afr_sumstats_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
amr_sumstats_file <- paste0("~/ukb_finemap/tmp_sumstats/AMR/", protein_id, ".txt")
write.table(amr_sumstats_filtered, amr_sumstats_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
eas_sumstats_file <- paste0("~/ukb_finemap/tmp_sumstats/EAS/", protein_id, ".txt")
write.table(eas_sumstats_filtered, eas_sumstats_file, sep= "\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
```

**04susier.R**
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(argparse))
suppressPackageStartupMessages(library(susieR))

#setup command line argument parsing
parser <- ArgumentParser()
parser$add_argument("--oid", required=TRUE, help="OID number (e.g., OID20089)")
args <- parser$parse_args()

protein_id <- args$oid
output_prefix <- paste0("~/ukb_finemap/susier_output/", protein_id)

sumstats_file <- fread(paste0("~/ukb_finemap/tmp_sumstats/META/", protein_id, ".txt"))
ld_file <- as.matrix(fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/META/", protein_id, "_common.txt"), header=TRUE))
#ld_file <- as.matrix(fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/META/", protein_id, "_common_fixed.txt"), header=TRUE))
common_snps <- fread(paste0("~/ukb_finemap/1kg/tmp_ld_files/common_snp_lists/", protein_id, ".txt"), header=FALSE)

zscores <- as.vector(sumstats_file$Z)

fitted_model <- tryCatch({
  susie_rss(
    z = zscores, #p-vector of z-scores
    R = ld_file, #p by p correlation matrix
    n = 34477, #sample size
    var_y = NULL,
    L = 10, #not default
    lambda = 0,
    maf = NULL,
    maf_thresh = 0,
    z_ld_weight = 0,
    prior_variance = NULL,
    scaled_prior_variance = 0.2,
    residual_variance = NULL,
    prior_weights = NULL,
    null_weight = NULL, #0 is default
    standardize = TRUE,
    intercept_value = 0,
    estimate_residual_variance = TRUE, #TRUE in MESA
    estimate_residual_method = c("MoM", "MLE"),
    estimate_prior_variance = TRUE,
    estimate_prior_method = "EM",
    unmappable_effects = c("none", "inf"),
    check_null_threshold = 0,
    prior_tol = 1e-09,
    residual_variance_lowerbound = 0,
    residual_variance_upperbound = Inf,
    model_init = NULL,
    coverage = 0.95,
    min_abs_corr = 0.5,
    max_iter = 100,
    tol = 0.001,
    convergence_method = c("elbo", "pip"),
    verbose = FALSE,
    track_fit = FALSE,
    check_input = FALSE,
    check_prior = TRUE,
    check_R = TRUE,
    check_z = FALSE,
    n_purity = 100,
    r_tol = 1e-08,
    refine = FALSE)
}, error = function(e) {
  stop(paste("Error in SuSiE analysis:", e$message))
})

#get PIPs
snp_names <- common_snps$V1
snp_pips <- data.frame(
  SNP = snp_names,
  PIP = fitted_model$pip
)
snp_pips <- snp_pips[order(-snp_pips$PIP), ]

#get credible sets
cs <- tryCatch({
  susie_get_cs(fitted_model, 
               X = NULL,
               Xcorr=ld_file,
               coverage = 0.95, 
               min_abs_corr = 0.5,
               dedup = TRUE,
               squared = FALSE)
}, error = function(e) {
  cat("Error getting credible sets:", conditionMessage(e), "\n")
  return(list())
})

#initialize empty results dataframe
cs_results <- data.frame(
  SNP = character(),
  CredibleSet = integer(),
  PIP = numeric(),
  stringsAsFactors = FALSE
)

#check which credible set structure we have and process accordingly
if (!is.null(cs$sets) && length(cs$sets) > 0) {
  # 'sets' structure
  cat("Processing credible sets from 'sets' structure...\n")
  
  for (i in seq_along(cs$sets)) {
    indices <- cs$sets[[i]]
    temp_results <- data.frame(
      SNP = snp_names[indices],
      CredibleSet = i,
      PIP = fitted_model$pip[indices],
      stringsAsFactors = FALSE
    )
    cs_results <- rbind(cs_results, temp_results)
  }
  for (i in seq_along(cs$sets)) {
    indices <- cs$sets[[i]]
    temp_results <- data.frame(
      SNP = snp_names[indices],
      CredibleSet = i,
      PIP = fitted_model$pip[indices],
      ComponentPIP = fitted_model$alpha[i, indices],
      stringsAsFactors = FALSE
    )
    cs_results <- rbind(cs_results, temp_results)
  }
} else {
  #try to find individual L sets (L1, L2, etc.)
  cat("Looking for individual L credible sets...\n")
  
  #first, check if they're directly in cs
  l_pattern <- "^L[0-9]+$"
  l_names <- names(cs)[grep(l_pattern, names(cs))]
  
  if (length(l_names) > 0) {
    for (l_name in l_names) {
      set_number <- as.integer(gsub("L", "", l_name))
      indices <- cs[[l_name]]
      if (length(indices) > 0) {
        temp_results <- data.frame(
          SNP = snp_names[indices],
          CredibleSet = set_number,
          PIP = fitted_model$pip[indices],
          stringsAsFactors = FALSE
        )
        cs_results <- rbind(cs_results, temp_results)
      }
    }
  } else if (!is.null(cs$cs)) {
    #if neither of the above, check if cs$cs exists
    l_names <- names(cs$cs)[grep(l_pattern, names(cs$cs))]
    
    if (length(l_names) > 0) {
      for (l_name in l_names) {
        set_number <- as.integer(gsub("L", "", l_name))
        indices <- cs$cs[[l_name]]
        if (length(indices) > 0) {
          temp_results <- data.frame(
            SNP = snp_names[indices],
            CredibleSet = set_number,
            PIP = fitted_model$pip[indices],
            stringsAsFactors = FALSE
          )
          cs_results <- rbind(cs_results, temp_results)
        }
      }
    }
  }
}


#sort and save if we have results
if (nrow(cs_results) > 0) {
  # Sort by credible set and PIP
  cs_results <- cs_results[order(cs_results$CredibleSet, -cs_results$PIP), ]
  
  #save to TSV
  write.table(cs_results, 
              file = paste0(output_prefix, "_susie_cis_cs.tsv"),
              sep = "\t", 
              quote = FALSE, 
              row.names = FALSE)
  
  cat("Credible sets saved to", paste0(output_prefix, "_susie_cis_cs.tsv"), "\n")
} else {
  cat("No credible sets found in any format.\n")
}
```

**05multi_susie.py**
```{python}
import numpy as np
import pandas as pd
import sys
import os
import argparse
import MultiSuSiE

#argparse
def check_arg(args=None):
    parser = argparse.ArgumentParser(description="Fine-map with multi-SuSiE for all proteins")
    parser.add_argument('--oid', help='OlinkID', required=True)
    return parser.parse_args(args)

args = check_arg(sys.argv[1:])
protein_id = args.oid

ancestries = ['EUR', 'AFR', 'AMR', 'EAS']

#build ld list with 0.02 ridge normalization
def build_rlist(protein_id):
    R_list = []
    for ancestry in ancestries:
        print(ancestry, "\n")
        try:
            tmp_ld_file = f'~/ukb_finemap/1kg/tmp_ld_files/{ancestry}/{protein_id}_common.txt'
            ld_matrix = pd.read_csv(tmp_ld_file, sep='\t', header=0).values.astype(np.float64)

            # Fixed LD regularization lambda = 0.02
            # R_reg = R * (1 - lambda) + lambda * I
            # Consistent with SuShiE --ld-adjust 0.02
            lambda_val = 0.02
            ld_matrix = ld_matrix * (1 - lambda_val) + lambda_val * np.eye(ld_matrix.shape[0])
            print(f"loaded and regularized LD matrix for {protein_id} {ancestry} (lambda = {lambda_val})")

            R_list.append(ld_matrix.astype(np.float32))
        except Exception as e:
            print(f"Error processing R_list {protein_id} {ancestry}: {e}")
            return None
    return R_list

#build z score lists
def build_zlist(protein_id):
    z_list = []
    for ancestry in ancestries:
        try:
            file_path = f'~/ukb_finemap/tmp_sumstats/{ancestry}/{protein_id}.txt'
            sumstats = pd.read_csv(file_path, sep='\t')
            # Convert to numpy array for low_memory_mode compatibility
            z_list.append(np.array(sumstats['Z'].tolist(), dtype=np.float32))
        except Exception as e:
            print(f"Error processing z_list {protein_id} {ancestry}: {e}")
            return None
    return z_list

#build snp list
def build_snp_list(protein_id):
    try:
        snp_path = f'~/ukb_finemap/1kg/tmp_ld_files/common_snp_lists/{protein_id}.txt'
        common_snp_ids = pd.read_csv(snp_path, header=None)[0].tolist()
    except Exception as e:
        print(f"Error processing snp_list {protein_id}: {e}")
        return None
    return common_snp_ids

#main
R_list = build_rlist(protein_id)
if R_list is None:
    print(f"NO_CREDIBLE_SETS: Failed to build LD list for {protein_id}")
    sys.exit(0)

z_list = build_zlist(protein_id)
if z_list is None:
    print(f"NO_CREDIBLE_SETS: Failed to build z_list for {protein_id}")
    sys.exit(0)

common_snp_ids = build_snp_list(protein_id)
if common_snp_ids is None:
    print(f"NO_CREDIBLE_SETS: Failed to build snp_list for {protein_id}")
    sys.exit(0)

#run MultiSuSiE
try:
    ss_fit = MultiSuSiE.multisusie_rss(
        z_list = z_list,
        R_list = R_list,
        varY_list = None,
        rho = np.array([[1.0,  0.75, 0.75, 0.75],
                        [0.75, 1.0,  0.75, 0.75],
                        [0.75, 0.75, 1.0,  0.75],
                        [0.75, 0.75, 0.75, 1.0]]),
        population_sizes = [33187, 931, 97, 262],
        L = 10,
        scaled_prior_variance = 0.2,
        prior_weights = None,
        standardize = False,
        pop_spec_standardization = True,
        estimate_residual_variance = True,
        estimate_prior_variance = True,
        estimate_prior_method = 'early_EM',
        pop_spec_effect_priors = True,
        iter_before_zeroing_effects = 5,
        prior_tol = 1e-9,
        max_iter = 100,
        tol = 0.001,
        verbose = False,
        coverage = 0.95,
        min_abs_corr = 0.5,
        float_type = np.float32,
        low_memory_mode = True,
        recover_R = False,
        single_population_mac_thresh = 0,
        mac_list = None,
        multi_population_maf_thresh = 0,
        maf_list = None,
        variant_ids = None
    )
except Exception as e:
    print(f"NO_CREDIBLE_SETS: MultiSuSiE failed for {protein_id}: {e}")
    sys.exit(0)

#extract credible sets
cs_results = []

if hasattr(ss_fit, 'sets') and ss_fit.sets[0]:
    credible_sets  = ss_fit.sets[0]
    purity         = ss_fit.sets[1]
    coverage       = ss_fit.sets[2]
    passed_filter  = ss_fit.sets[3]

    has_valid_cs = False
    for cs_idx, credible_set in enumerate(credible_sets):
        if passed_filter[cs_idx] and len(credible_set) > 0:
            has_valid_cs = True
            for variant_idx in credible_set:
                actual_snp_id = (common_snp_ids[variant_idx]
                                 if variant_idx < len(common_snp_ids)
                                 else f"variant_{variant_idx}")
                cs_results.append({
                    'OlinkID':     protein_id,
                    'SNP':         actual_snp_id,
                    'Index':       variant_idx,
                    'CredibleSet': cs_idx + 1,
                    'PIP':         ss_fit.pip[variant_idx],
                    'Purity':      purity[cs_idx],
                    'Coverage':    coverage[cs_idx]
                })

    if has_valid_cs:
        cs_results_df = pd.DataFrame(cs_results)
        cs_results_df.to_csv(
            f'~/ukb_finemap/multi_output_rho/{protein_id}_credible_sets.tsv',
            sep="\t", index=False
        )
        print(f"SUCCESS: Multi-SuSiE found {len(set([r['CredibleSet'] for r in cs_results]))} credible sets for {protein_id}")
    else:
        print(f"NO_CREDIBLE_SETS: Multi-SuSiE found no valid credible sets for {protein_id}")
else:
    print(f"NO_CREDIBLE_SETS: Multi-SuSiE did not identify any credible sets for {protein_id}")
```

**06susiex_ld_calc.py**
```{python}
import numpy as np
import pandas as pd
import struct
import os
import sys
import argparse

#set up argparse
def check_arg(args=None):
    parser = argparse.ArgumentParser(description="Fine-map with SuSiEx")
    parser.add_argument('--oid', help='OlinkID', required='True')
    return parser.parse_args(args)

args = check_arg(sys.argv[1:])

protein_id = args.oid

ancestries = ['EUR', 'AFR', 'AMR', 'EAS']

for ancestry in ancestries:
    ld_in = f'~/ukb_finemap/1kg/tmp_ld_files/{ancestry}/{protein_id}_common.txt'
    out_prefix = f'~/ukb_finemap/1kg/tmp_ld_files/{ancestry}/{protein_id}_susiex'
    os.makedirs(os.path.dirname(out_prefix), exist_ok=True)

    print(f"Processing susiex ld for {ancestry}...")

    #load LD matrix — header row contains SNP IDs (with chr prefix), no row index
    ld_df = pd.read_csv(ld_in, sep='\t', header=0, index_col=None)
    snp_ids_chr = list(ld_df.columns)               # e.g. 'chr4:163197091:A:G'
    snp_ids = [s.replace('chr', '') for s in snp_ids_chr]  # e.g. '4:163197091:A:G'
    ld_matrix = ld_df.values.astype(np.float32)
    n_snps = ld_matrix.shape[0]

    #write .ld.bin
    with open(f'{out_prefix}.ld.bin', 'wb') as f:
        f.write(ld_matrix.astype(np.float32).tobytes())

    #write _ref.bim (no header, tab-separated: CHR SNP CM BP A1 A2)
    #write _ref.bim
    rows = []
    for snp in snp_ids:
        parts = snp.split(':')
        chrom, pos, a1, a2 = parts[0].replace("X", "23"), parts[1], parts[2], parts[3]
        rows.append([chrom, snp, 0, pos, a1, a2])
    bim_df = pd.DataFrame(rows, columns=['CHR', 'SNP', 'CM', 'BP', 'A1', 'A2'])
    bim_df = bim_df.assign(CHR=bim_df['CHR'].astype(str).str.replace("X", "23", regex=False))
    bim_df.to_csv(f'{out_prefix}_ref.bim', sep='\t', index=False, header=False)

    #write _frq.frq (CHR SNP A1 A2 MAF NCHROBS)
    try:
        sumstats = pd.read_csv(
            f'~/ukb_finemap/tmp_sumstats/{ancestry}/{protein_id}.txt',
            sep='\t'
        )
        #sumstats SNP IDs are in ID column, no chr prefix — matches snp_ids
        sumstats = sumstats.set_index('ID')

        #convert A1FREQ to MAF (minor allele freq = min of freq and 1-freq)
        sumstats['MAF'] = sumstats['A1FREQ'].apply(lambda x: min(x, 1 - x))

        #estimate N per variant from SE: N ~ 1/SE^2
        sumstats['N_est'] = (1.0 / sumstats['SE'] ** 2).round().astype(int)

        frq_rows = []
        n_matched = 0
        for snp in snp_ids:
            parts = snp.split(':')
            chrom, a1, a2 = parts[0].replace("X", "23"), parts[2], parts[3]
            if snp in sumstats.index:
                maf = sumstats.loc[snp, 'MAF']
                nchrobs = sumstats.loc[snp, 'N_est'] * 2
                n_matched += 1
            else:
                maf = 'NA'
                nchrobs = 'NA'
            frq_rows.append([chrom, snp, a1, a2, maf, nchrobs])

        frq_df = pd.DataFrame(frq_rows, columns=['CHR', 'SNP', 'A1', 'A2', 'MAF', 'NCHROBS'])
        frq_df = frq_df.assign(CHR=frq_df['CHR'].astype(str).str.replace("X", "23", regex=False))
        frq_df.to_csv(f'{out_prefix}_frq.frq', sep='\t', index=False)

        if n_matched < n_snps:
            print(f"  WARNING: {n_snps - n_matched} SNPs in LD matrix had no sumstats match")

    except Exception as e:
        print(f"  ERROR building _frq.frq for {ancestry}: {e}")

    print(f"  Done: {out_prefix}{{.ld.bin, _ref.bim, _frq.frq}}")
```

**Combine Files and Merge with MESA Fine-Mapping Output**
```{r}
library(dplyr)
library(data.table)
library(stringr)
library(tidyr)

#SuSiE R
root_dir <- "~/ukb_finemap/susier_output"
susier_files <- list.files(path = root_dir, pattern = "*_susie_cis_cs.tsv", recursive = TRUE, full.names = TRUE)
susier_data <- list()
for (i in 1:length(susier_files)) {
  file <- susier_files[i]
  olink_id <- str_extract(file, "OID\\d+")
  data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  data$OlinkID <- olink_id
  colnames(data) <- c("SNP_UKB", "susieR_CS_UKB", "susieR_PIP_UKB", "OlinkID")
  data <- data %>% select(OlinkID, SNP_UKB, susieR_CS_UKB, susieR_PIP_UKB)
  susier_data[[i]] <- data
}
susier_combined <- do.call(rbind, susier_data)
susier_combined <- susier_combined %>%
  filter(!is.na(SNP_UKB)) %>%
  group_by(OlinkID, susieR_CS_UKB) %>%
  mutate(max_pip_in_cs = max(susieR_PIP_UKB)) %>%
  ungroup() %>%
  group_by(OlinkID, SNP_UKB) %>%
  slice_max(max_pip_in_cs, n = 1, with_ties = FALSE) %>%
  ungroup()

#SuShiE
root_dir <- "~/ukb_finemap/sushie_output"
sushie_files <- list.files(path = root_dir, pattern = "*.sushie.cs.tsv", recursive = TRUE, full.names = TRUE)
sushie_data <- list()
for (i in 1:length(sushie_files)) {
  file <- sushie_files[i]
  data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  data_filtered <- data %>% select(trait, snp, CSIndex, pip_cs)
  colnames(data_filtered) <- c("OlinkID", "SNP_UKB", "sushie_CS_UKB", "sushie_PIP_UKB")
  sushie_data[[i]] <- data_filtered
}
sushie_combined <- do.call(rbind, sushie_data)
sushie_combined <- sushie_combined %>%
  filter(!is.na(SNP_UKB)) %>%
  group_by(OlinkID, sushie_CS_UKB) %>%
  mutate(max_pip_in_cs = max(sushie_PIP_UKB)) %>%
  ungroup() %>%
  group_by(OlinkID, SNP_UKB) %>%
  slice_max(max_pip_in_cs, n = 1, with_ties = FALSE) %>%
  ungroup()

#MultiSuSiE
root_dir <- "~/ukb_finemap/multi_output_rho"
multi_files <- list.files(path = root_dir, pattern = "*_credible_sets.tsv", recursive = TRUE, full.names = TRUE)
multi_data <- list()
for (i in 1:length(multi_files)) {
  file <- multi_files[i]
  data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  data_filtered <- data %>% select(OlinkID, SNP, CredibleSet, PIP)
  colnames(data_filtered) <- c("OlinkID", "SNP_UKB", "multi_CS_UKB", "multi_PIP_UKB")
  multi_data[[i]] <- data_filtered
}
multi_combined <- do.call(rbind, multi_data)
multi_combined <- multi_combined %>%
  filter(!is.na(SNP_UKB)) %>%
  group_by(OlinkID, multi_CS_UKB) %>%
  mutate(max_pip_in_cs = max(multi_PIP_UKB)) %>%
  ungroup() %>%
  group_by(OlinkID, SNP_UKB) %>%
  slice_max(max_pip_in_cs, n = 1, with_ties = FALSE) %>%
  ungroup()

#SuSiE X
root_dir <- "~/ukb_finemap/susiex_output"
susiex_files <- list.files(path = root_dir, pattern = "*.cs", recursive = TRUE, full.names = TRUE)
susiex_data <- list()
for (i in 1:length(susiex_files)) {
  file <- susiex_files[i]
  olink_id <- str_extract(file, "OID\\d+")
  data <- read.table(file, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  if (is.null(data) || nrow(data) == 0) next
  data$OlinkID <- olink_id
  data_filtered <- data %>% select(OlinkID, SNP, CS_ID, CS_PIP)
  colnames(data_filtered) <- c("OlinkID", "SNP_UKB", "susieX_CS_UKB", "susieX_PIP_UKB")
  susiex_data[[i]] <- data_filtered
}
susiex_combined <- do.call(rbind, susiex_data)
susiex_combined <- susiex_combined %>%
  filter(!is.na(SNP_UKB)) %>%
  group_by(OlinkID, susieX_CS_UKB) %>%
  mutate(max_pip_in_cs = max(susieX_PIP_UKB)) %>%
  ungroup() %>%
  group_by(OlinkID, SNP_UKB) %>%
  slice_max(max_pip_in_cs, n = 1, with_ties = FALSE) %>%
  ungroup()

#combined
combined_results <- susier_combined %>%
  full_join(sushie_combined, by = c("OlinkID", "SNP_UKB")) %>%
  full_join(multi_combined, by = c("OlinkID", "SNP_UKB")) %>%
  full_join(susiex_combined, by = c("OlinkID", "SNP_UKB"))
write.table(combined_results, "~/ukb_finemap/combined_UKB.txt", col.names = TRUE, row.names = FALSE, quote = FALSE, sep = "\t")

combined_results_ukb <- combined_results %>%
  separate(SNP_UKB, into = c("CHR", "POS_hg37", "REF", "ALT"), sep = ":", remove = FALSE) %>%
  mutate(REF = NULL, ALT = NULL)

#build 37 to 38
coords <- fread("~/ukb_finemap/snp_coords_hg37_to_hg38.tsv")
colnames(coords) <- c("CHR", "POS", "POS_hg37")
coords$CHR <- gsub("^chr", "", coords$CHR)
coords$CHR <- as.character(coords$CHR)
coords <- coords %>%
  mutate(CHR = ifelse(CHR == 23, "X", as.character(CHR)))
coords$POS_hg37 <- as.character(coords$POS_hg37)
coords$POS <- as.character(coords$POS)
combined_results_ukb_w_coords <- left_join(combined_results_ukb, coords, by=c("CHR", "POS_hg37"))

#MESA data
mesa_data <- fread("~/compare_susie/meta_cis_tensor_vs_susier_vs_sushie_vs_multi_vs_susiex.tsv")
colnames(mesa_data) <- c("OlinkID", "SNP_MESA", "tensor_CS_MESA", "tensor_PIP_MESA", "susieR_CS_MESA", "susieR_PIP_MESA", "sushie_CS_MESA", "sushie_PIP_MESA", "multi_CS_MESA", "multi_PIP_MESA", "susieX_CS_MESA", "susieX_PIP_MESA")
mesa_data$SNP_MESA <- gsub("^chr", "", mesa_data$SNP_MESA)
mesa_data$tensor_CS_MESA <- gsub("^L", "", mesa_data$tensor_CS_MESA)
#mesa_data <- mesa_data %>% filter(OlinkID %in% combined_results$OlinkID)
combined_results_mesa <- mesa_data %>%
  separate(SNP_MESA, into = c("CHR", "POS", "REF", "ALT"), sep = ":", remove = FALSE) %>%
  mutate(REF = NULL, ALT = NULL)

shared_proteins <- intersect(
  unique(combined_results_ukb_w_coords$OlinkID),
  unique(combined_results_mesa$OlinkID)
)

ukb_mesa_data <- full_join(combined_results_ukb_w_coords, combined_results_mesa, by=c("OlinkID", "CHR", "POS")) %>%
  select(OlinkID, CHR, POS, SNP_MESA, SNP_UKB, tensor_CS_MESA, susieR_CS_MESA, susieR_CS_UKB, tensor_PIP_MESA, susieR_PIP_MESA, susieR_PIP_UKB, sushie_CS_MESA, sushie_CS_UKB, sushie_PIP_MESA, sushie_PIP_UKB, multi_CS_MESA, multi_CS_UKB, multi_PIP_MESA, multi_PIP_UKB, susieX_CS_MESA, susieX_CS_UKB, susieX_PIP_MESA, susieX_PIP_UKB) %>%
  mutate(
    CHR = ifelse(CHR == 23, "X", as.character(CHR)),
    CHR = factor(CHR, levels = c(as.character(1:22), "X")),
    POS = as.numeric(POS)
  ) %>%
  arrange(OlinkID, CHR, POS)

dupes <- ukb_mesa_data[duplicated(ukb_mesa_data[c("OlinkID", "SNP_MESA", "SNP_UKB")]) | duplicated(ukb_mesa_data[c("OlinkID", "SNP_MESA", "SNP_UKB")], fromLast = TRUE), ]
dupes[order(dupes$OlinkID, dupes$SNP_MESA, dupes$SNP_UKB), ]

ukb_mesa_data_filtered <- ukb_mesa_data %>%
  filter(OlinkID %in% shared_proteins)

write.table(ukb_mesa_data_filtered, "~/ukb_finemap/combined_MESA_UKB.txt", col.names = TRUE, row.names = FALSE, quote = FALSE, sep = "\t")

#SNP/Protein Pairs
ukb_mesa_overlap <- ukb_mesa_data %>%
  filter(
    if_any(ends_with("_CS_MESA"), ~ !is.na(.) & . != 0) &
      if_any(ends_with("_CS_UKB"),  ~ !is.na(.) & . != 0)
  ) %>%
  select(OlinkID, CHR, POS, SNP_MESA, SNP_UKB, tensor_CS_MESA, susieR_CS_MESA, susieR_CS_UKB, tensor_PIP_MESA, susieR_PIP_MESA, susieR_PIP_UKB, sushie_CS_MESA, sushie_CS_UKB, sushie_PIP_MESA, sushie_PIP_UKB, multi_CS_MESA, multi_CS_UKB, multi_PIP_MESA, multi_PIP_UKB, susieX_CS_MESA, susieX_CS_UKB, susieX_PIP_MESA, susieX_PIP_UKB)
write.table(ukb_mesa_overlap, "~/ukb_finemap/combined_MESA_UKB_replicates.txt", col.names = TRUE, row.names = FALSE, quote = FALSE, sep = "\t")
```

**Calculate Precision, Recall, and F1 for all combinations, excluding MESA SusieR**
```{r}
library(data.table)
library(dplyr)
library(tidyr)

#all SNPs tested
all_mesa_snps <- fread("~/ukb_finemap/overlap_snps/mesa_overlap.txt")
all_ukb_snps <- fread("~/ukb_finemap/overlap_snps/ukb_overlap.txt")

#cross-method overlapping CS files
all_mesa_snps_susier$POS_hg38 <- as.integer(all_mesa_snps_susier$POS_hg38)
all_mesa_snps_susier$OlinkID <- all_mesa_snps_susier$phenotype_id

#finemapped SNPs
combined_finemap <- fread("~/ukb_finemap/combined_MESA_UKB.txt")
input_oids <- fread("~/tensor_susie/all_race/0.05cis_locations.tsv") %>%
  select(phenotype_id)

#filter for 1534 proteins with pQTLs FDR < 0.05
combined_finemap <- combined_finemap %>%
  filter(OlinkID %in% input_oids$phenotype_id)

#ran with MESA susieR SNPs only when susieR is evaluated
overlap_snps_tested <- all_mesa_snps %>%
  inner_join(all_ukb_snps, by = c("OlinkID", "CHR", "POS_hg38"))

mesa_methods <- c("sushie", "multi", "susieX")
ukb_methods  <- c("susieR", "sushie", "multi", "susieX")

for (m_ukb in ukb_methods) {
  for (m_mesa in mesa_methods) {
    
    mesa_CS_col  <- paste0(m_mesa, "_CS_MESA")
    ukb_CS_col   <- paste0(m_ukb,  "_CS_UKB")
    mesa_PIP_col <- paste0(m_mesa, "_PIP_MESA")
    ukb_PIP_col  <- paste0(m_ukb,  "_PIP_UKB")
    
    proteins_both <- intersect(
      combined_finemap %>% filter(!is.na(.data[[mesa_CS_col]])) %>% pull(OlinkID) %>% unique(),
      combined_finemap %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% pull(OlinkID) %>% unique()
    )
    
    cross_data <- combined_finemap %>%
      filter(OlinkID %in% proteins_both) %>%
      select(OlinkID, CHR, POS, SNP_MESA, SNP_UKB,
             all_of(c(mesa_CS_col, ukb_CS_col, mesa_PIP_col, ukb_PIP_col)))
    
    cross_intersect <- cross_data %>%
      group_by(OlinkID, CHR, POS) %>%
      filter(!is.na(.data[[mesa_CS_col]]) & !is.na(.data[[ukb_CS_col]]))
    
    cross_UKB <- cross_data %>%
      select(OlinkID, CHR, POS, SNP_UKB, all_of(c(ukb_CS_col, ukb_PIP_col))) %>%
      filter(!is.na(.data[[ukb_CS_col]])) %>%
      group_by(OlinkID, .data[[ukb_CS_col]]) %>%
      filter(any(paste(OlinkID, CHR, POS, .data[[ukb_CS_col]]) %in%
                   paste(cross_intersect$OlinkID, cross_intersect$CHR, cross_intersect$POS, cross_intersect[[ukb_CS_col]]))) %>%
      ungroup()
    
    cross_MESA <- cross_data %>%
      select(OlinkID, CHR, POS, SNP_MESA, all_of(c(mesa_CS_col, mesa_PIP_col))) %>%
      filter(!is.na(.data[[mesa_CS_col]])) %>%
      group_by(OlinkID, .data[[mesa_CS_col]]) %>%
      filter(any(paste(OlinkID, CHR, POS, .data[[mesa_CS_col]]) %in%
                   paste(cross_intersect$OlinkID, cross_intersect$CHR, cross_intersect$POS, cross_intersect[[mesa_CS_col]]))) %>%
      ungroup()
    
    write.table(cross_MESA, paste0("~/ukb_finemap/overlap_snps/", m_mesa, "_MESA_", m_ukb, "_UKB_overlapping_cs.txt"),     quote=FALSE, row.names=FALSE, sep="\t")
    write.table(cross_UKB,  paste0("~/ukb_finemap/overlap_snps/", m_mesa, "_MESA_", m_ukb, "_UKB_overlapping_cs_UKB.txt"), quote=FALSE, row.names=FALSE, sep="\t")
    
    assign(paste0(m_mesa, "_", m_ukb, "_cross_data"),      cross_data)
    assign(paste0(m_mesa, "_", m_ukb, "_proteins_both"),   proteins_both)
  }
}

#contingency tables
for (m_ukb in ukb_methods) {
  for (m_mesa in mesa_methods) {
    
    cat("\n=== MESA:", m_mesa, "vs UKB:", m_ukb, "===\n")
    
    mesa_CS_col <- paste0(m_mesa, "_CS_MESA")
    ukb_CS_col  <- paste0(m_ukb,  "_CS_UKB")
    
    proteins_both <- get(paste0(m_mesa, "_", m_ukb, "_proteins_both"))
    pair_data     <- get(paste0(m_mesa, "_", m_ukb, "_cross_data"))
    overlap       <- overlap_snps_tested %>% filter(OlinkID %in% proteins_both)
    
    TP <- overlap %>%
      inner_join(pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      inner_join(pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    FP <- overlap %>%
      inner_join(pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      anti_join( pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    FN <- overlap %>%
      inner_join(pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      anti_join( pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    TN <- overlap %>%
      anti_join(pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      anti_join(pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    contingency <- data.frame(c(TP, FN), c(FP, TN))
    colnames(contingency) <- c("SNP_in_CS", "SNP_out_CS")
    rownames(contingency) <- c("SNP_in_CS", "SNP_out_CS")
    print(contingency)
    
    precision   <- TP / (TP + FP)
    sensitivity <- TP / (TP + FN)
    specificity <- TN / (TN + FP)
    f1          <- 2 * TP / (2 * TP + FP + FN)
    accuracy    <- (TP + TN) / (TP + TN + FP + FN)
    
    cat(sprintf("\n  TP : %d  FP : %d  FN : %d  TN : %d\n", TP, FP, FN, TN))
    cat(sprintf("  Sensitivity [TP/(TP+FN)]      : %.3f\n", sensitivity))
    cat(sprintf("  Specificity [TN/(TN+FP)]      : %.3f\n", specificity))
    cat(sprintf("  Precision   [TP/(TP+FP)]      : %.3f\n", precision))
    cat(sprintf("  F1          [2TP/(2TP+FP+FN)] : %.3f\n", f1))
    cat(sprintf("  Accuracy    [(TP+TN)/total]   : %.3f\n", accuracy))
  }
}
```

**Calculate Precision, Recall, and F1 for all combinations with MESA SusieR**
```{r}
library(data.table)
library(dplyr)
library(tidyr)

#all SNPs tested
all_mesa_snps <- fread("~/ukb_finemap/overlap_snps/mesa_overlap.txt")
all_ukb_snps <- fread("~/ukb_finemap/overlap_snps/ukb_overlap.txt")

all_mesa_snps_susier <- fread("~/qtl_mapping/all_race/tensorqtl/nominal/combined_nominal_qtls.txt")
all_mesa_snps_susier <- all_mesa_snps_susier %>%
  separate(variant_id, into = c("CHR", "POS_hg38", "REF", "ALT"), sep = ":", remove = FALSE)
all_mesa_snps_susier <- all_mesa_snps_susier %>%
  mutate(
    CHR = as.character(gsub("^chr", "", all_mesa_snps_susier$CHR)),
    POS_hg38 = all_mesa_snps_susier$POS)

#cross-method overlapping CS files
all_mesa_snps_susier$POS_hg38 <- as.integer(all_mesa_snps_susier$POS_hg38)
all_mesa_snps_susier$OlinkID <- all_mesa_snps_susier$phenotype_id

#finemapped SNPs
combined_finemap <- fread("~/ukb_finemap/combined_MESA_UKB.txt")
input_oids <- fread("~/tensor_susie/all_race/0.05cis_locations.tsv") %>%
  select(phenotype_id)

#filter for 1534 proteins with pQTLs FDR < 0.05
combined_finemap <- combined_finemap %>%
  filter(OlinkID %in% input_oids$phenotype_id)

#ran with MESA susieR SNPs only when susieR is evaluated
overlap_snps_tested <- all_mesa_snps %>%
  inner_join(all_ukb_snps, by = c("OlinkID", "CHR", "POS_hg38")) %>%
  inner_join(all_mesa_snps_susier, by = c("OlinkID", "CHR", "POS_hg38"))

mesa_methods <- c("susieR")
ukb_methods  <- c("susieR", "sushie", "multi", "susieX")

for (m_ukb in ukb_methods) {
  for (m_mesa in mesa_methods) {
    
    mesa_CS_col  <- paste0(m_mesa, "_CS_MESA")
    ukb_CS_col   <- paste0(m_ukb,  "_CS_UKB")
    mesa_PIP_col <- paste0(m_mesa, "_PIP_MESA")
    ukb_PIP_col  <- paste0(m_ukb,  "_PIP_UKB")
    
    proteins_both <- intersect(
      combined_finemap %>% filter(!is.na(.data[[mesa_CS_col]])) %>% pull(OlinkID) %>% unique(),
      combined_finemap %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% pull(OlinkID) %>% unique()
    )
    
    cross_data <- combined_finemap %>%
      filter(OlinkID %in% proteins_both) %>%
      select(OlinkID, CHR, POS, SNP_MESA, SNP_UKB,
             all_of(c(mesa_CS_col, ukb_CS_col, mesa_PIP_col, ukb_PIP_col)))
    
    cross_intersect <- cross_data %>%
      group_by(OlinkID, CHR, POS) %>%
      filter(!is.na(.data[[mesa_CS_col]]) & !is.na(.data[[ukb_CS_col]]))
    
    cross_UKB <- cross_data %>%
      select(OlinkID, CHR, POS, SNP_UKB, all_of(c(ukb_CS_col, ukb_PIP_col))) %>%
      filter(!is.na(.data[[ukb_CS_col]])) %>%
      group_by(OlinkID, .data[[ukb_CS_col]]) %>%
      filter(any(paste(OlinkID, CHR, POS, .data[[ukb_CS_col]]) %in%
                   paste(cross_intersect$OlinkID, cross_intersect$CHR, cross_intersect$POS, cross_intersect[[ukb_CS_col]]))) %>%
      ungroup()
    
    cross_MESA <- cross_data %>%
      select(OlinkID, CHR, POS, SNP_MESA, all_of(c(mesa_CS_col, mesa_PIP_col))) %>%
      filter(!is.na(.data[[mesa_CS_col]])) %>%
      group_by(OlinkID, .data[[mesa_CS_col]]) %>%
      filter(any(paste(OlinkID, CHR, POS, .data[[mesa_CS_col]]) %in%
                   paste(cross_intersect$OlinkID, cross_intersect$CHR, cross_intersect$POS, cross_intersect[[mesa_CS_col]]))) %>%
      ungroup()
    
    write.table(cross_MESA, paste0("~/ukb_finemap/overlap_snps/", m_mesa, "_MESA_", m_ukb, "_UKB_overlapping_cs.txt"),     quote=FALSE, row.names=FALSE, sep="\t")
    write.table(cross_UKB,  paste0("~/ukb_finemap/overlap_snps/", m_mesa, "_MESA_", m_ukb, "_UKB_overlapping_cs_UKB.txt"), quote=FALSE, row.names=FALSE, sep="\t")
    
    assign(paste0(m_mesa, "_", m_ukb, "_cross_data"),      cross_data)
    assign(paste0(m_mesa, "_", m_ukb, "_proteins_both"),   proteins_both)
  }
}

#contingency tables
for (m_ukb in ukb_methods) {
  for (m_mesa in mesa_methods) {
    
    cat("\n=== MESA:", m_mesa, "vs UKB:", m_ukb, "===\n")
    
    mesa_CS_col <- paste0(m_mesa, "_CS_MESA")
    ukb_CS_col  <- paste0(m_ukb,  "_CS_UKB")
    
    proteins_both <- get(paste0(m_mesa, "_", m_ukb, "_proteins_both"))
    pair_data     <- get(paste0(m_mesa, "_", m_ukb, "_cross_data"))
    overlap       <- overlap_snps_tested %>% filter(OlinkID %in% proteins_both)
    
    TP <- overlap %>%
      inner_join(pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      inner_join(pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    FP <- overlap %>%
      inner_join(pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      anti_join( pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    FN <- overlap %>%
      inner_join(pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      anti_join( pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                 by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    TN <- overlap %>%
      anti_join(pair_data %>% filter(!is.na(.data[[mesa_CS_col]])) %>% select(OlinkID, CHR, POS),
                by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      anti_join(pair_data %>% filter(!is.na(.data[[ukb_CS_col]]))  %>% select(OlinkID, CHR, POS),
                by = c("OlinkID", "CHR", "POS_hg38" = "POS")) %>%
      nrow()
    
    contingency <- data.frame(c(TP, FN), c(FP, TN))
    colnames(contingency) <- c("SNP_in_CS", "SNP_out_CS")
    rownames(contingency) <- c("SNP_in_CS", "SNP_out_CS")
    print(contingency)
    
    precision   <- TP / (TP + FP)
    sensitivity <- TP / (TP + FN)
    specificity <- TN / (TN + FP)
    f1          <- 2 * TP / (2 * TP + FP + FN)
    accuracy    <- (TP + TN) / (TP + TN + FP + FN)
    
    cat(sprintf("\n  TP : %d  FP : %d  FN : %d  TN : %d\n", TP, FP, FN, TN))
    cat(sprintf("  Sensitivity [TP/(TP+FN)]      : %.3f\n", sensitivity))
    cat(sprintf("  Specificity [TN/(TN+FP)]      : %.3f\n", specificity))
    cat(sprintf("  Precision   [TP/(TP+FP)]      : %.3f\n", precision))
    cat(sprintf("  F1          [2TP/(2TP+FP+FN)] : %.3f\n", f1))
    cat(sprintf("  Accuracy    [(TP+TN)/total]   : %.3f\n", accuracy))
  }
}
```
