---
title: "Covariate Adjustment"
author: "Claudia Krueger"
date: "2024-12-15"
output: html_document
---
Prepare covariate files

**input tables**
```{r}
#read in protein expression matrix
olink <- fread("~/ancestry_specific_folder/post_qc_merge_olink.txt", header = TRUE)

#read in filtered demographic data
demographic <- fread("~/ancestry_specific_folder/protein_sidno_data", header = TRUE)

#read in genotype eigenvector matrix
pcs <- fread("~/ancestry_specific_folder/pca.eigenvec")
```

**format demographic data and extract necessary columns**
```{r}
#arrange demographic data by SMP ID number
ordered_demographic <- demographic %>% 
  mutate(ID_num = as.numeric(sub("SMP_", "", SampleID))) %>%  #extract numeric part
  arrange(ID_num) %>%  #arrange by the numeric part
  select(-ID_num)  #remove the temporary column

#extract sidno and smp columns
sidno_smp <- ordered_demographic[, c("SampleID", "sidno")]
sidno <- sidno_smp$sidno

#extract fid and smp columns
fid_smp <- demographic[, c("FID", "SampleID")]

#filter eigenvec file to keep only people/samples that are in the ancestry-specific filtered samples and order rows numerically
pcs_w_smps <- inner_join(pcs, fid_smp, by=c("FID"))
filtered_pcs <- pcs_w_smps %>% 
  select(SampleID, PC1, PC2, PC3, PC4, PC5, PC6, PC7, PC8, PC9, PC10)%>%
  mutate(ID_num = as.numeric(sub("SMP_", "", SampleID))) %>%  #extract numeric part
  arrange(ID_num) %>%  #arrange by the numeric part
  select(-ID_num)  #remove the temporary column

#merge genotype pcs and selected covariates into one table
all_covariates <- left_join(filtered_pcs, ordered_demographic, by=c("SampleID"))

#extract desired columns
selected_covariates <-  all_covariates %>% select(SampleID, PC1, PC2, PC3, PC4, PC5, PC6, PC7, PC8, PC9, PC10, race1c.x, gender1, AGE_AT_COLLECTION)

#remove columns with NA values
covariates_no_na <- na.omit(selected_covariates)
```

**use only proteins with matching sample IDs in demographic data**
```{r}
#identify the common SampleIDs
common_sample_ids <- intersect(covariates_no_na$SampleID, olink$SampleID)

#filter both tables to keep only rows with common SampleIDs
filtered_covariates <- covariates_no_na %>%
  filter(SampleID %in% common_sample_ids)

filtered_olink <- as.data.frame(olink) %>%
  filter(SampleID %in% common_sample_ids)

filtered_sidno <- sidno_smp %>%
  filter(SampleID %in% common_sample_ids)

filtered_pcs <- filtered_pcs %>%
  filter(SampleID %in% common_sample_ids)
```

**save new files**
```{r}
write.table(filtered_pcs, "~/ancestry_specific_folder/protein_adj_pca.eigenvec", quote=FALSE, sep="\t", row.names = FALSE, col.names =TRUE)
write.table(filtered_olink[,-1], "~/ancestry_specific_folder/covar_filtered_olink", quote=FALSE, sep="\t", row.names = FALSE, col.names =TRUE)
write.table(filtered_covariates[,-1], "~/ancestry_specific_folder/na_filtered_covariates", quote=FALSE, sep="\t", row.names = FALSE, col.names =TRUE)
write.table(filtered_sidno[,-1], "~/ancestry_specific_folder/covar_filtered_sidno", quote=FALSE, sep="\t", row.names = FALSE, col.names =TRUE)
```

**input files or use above variables**
```{r}
covariates <- fread("~/ancestry_specific_folder/na_filtered_covariates")
olink <- fread("~/ancestry_specific_folder/covar_filtered_olink")
sidno <- fread("~/ancestry_specific_folder/covar_filtered_sidno")
smp_lables <- fread("~/ancestry_specific_folder/protein_adj_pca.eigenvec")
smps <- smp_lables[,1]
```

**convert data types**
```{r}
#make variables factors
covariates$gender1 <- as.factor(covariates$gender1)
covariates$race1c.x <- as.factor(covariates$race1c.x)
covariates$COLLECTION_YEAR <- as.factor(covariates$COLLECTION_YEAR)
covariates$AGE_AT_COLLECTION <- as.factor(covariates$AGE_AT_COLLECTION)
sidno$sidno <- as.factor(sidno$sidno)
sidno <- sidno$sidno

#convert to data frame
olink <- as.data.frame(olink)
```

**perform covariate adjustment with linear mixed modeling (lmer/ranefmat) and linear modeling (lm/residmat)**
```{r}
#generate matrix
ranefmat <- matrix(NA, nrow = length(unique(sidno)), ncol = dim(olink)[2])
residmat <- matrix(NA,nrow=dim(olink)[1],ncol=dim(olink)[2])

#perform lmer
for(i in 1:dim(olink)[2]){
  fit <- lmer(olink[,i] ~ (1|sidno) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + gender1 + COLLECTION_YEAR + AGE_AT_COLLECTION, data = covariates, control = lmerControl(check.conv.singular = "ignore"))
  fitranef <- ranef(fit)
  fitresid <- resid(fit)
  ranefmat[,i] <- fitranef$sidno[,1]
  residmat[,i] <- fitresid
}

#label rows and columns of ranefmat
colnames(ranefmat) <-colnames(olink)
rownames(ranefmat) <- rownames(fitranef$sidno)
ranefmat <- data.frame(SampleID = rownames(ranefmat), ranefmat)
ranefmat$SampleID <- as.numeric(ranefmat$SampleID)

#label rows and columns of residmat
rownames(residmat) <- smps$SampleID
colnames(residmat) <- colnames(olink)
residmat <- data.frame(SampleID = rownames(residmat), residmat)

#read in table with sidno ids and corresponding fids
fids <- fread("~/ancestry_specific_folder/protein_sidno_data", header=TRUE)

#select columns
filtered_fids <- fids[ ,c("FID", "sidno")]

#merge by sidno
fids_w_ranefmat <- left_join(ranefmat, filtered_fids, by=c("SampleID" = "sidno"))

#move the last column to the first
fids_w_ranefmat <- fids_w_ranefmat %>%
  select(last_col(), everything())

#remove sidno column
fids_w_ranefmat <- fids_w_ranefmat[, -2]

#keep only unique rows
fids_w_ranefmat <- fids_w_ranefmat %>%
  distinct()

#save final files
write.table(fids_w_ranefmat, "~/ancestry_specific_folder/ranefmat", row.names=FALSE, col.names=TRUE, sep="\t", quote=FALSE)
write.table(residmat, "~/ancestry_specific_folder/residmat", row.names=FALSE, col.names=TRUE, sep="\t", quote=FALSE)
```

**edit final linear model table**
```{r}
olink_matrix <- fread("~/ancestry_specific_folder/post_qc_merge_olink.txt", header = TRUE)
residmat <- read.table("~/ancestry_specific_folder/residmat", header=TRUE)
demographic <- fread("~/ancestry_specific_folder/protein_sidno_data", header = TRUE)

#filter original Olink matrix and demographic data
filtered_olink <- olink_matrix %>%
  filter(olink_matrix$SampleID %in% (residmat$SampleID))

filtered_demographic <- demographic %>%
  filter(demographic$SampleID %in% (residmat$SampleID))

#isolate ID columns
fid_sidno_snp <- select(filtered_demographic, "FID", "sidno", "SampleID")

#add sidno ids based on corresponding smp
residmat_sidno <- left_join(fid_sidno_snp, residmat, by=c("SampleID"))
residmat_fid <- residmat_sidno[, -c(2, 3)]

olink_sidno <- left_join(fid_sidno_snp, filtered_olink, by=c("SampleID"))
olink_fid <- olink_sidno[, -c(2, 3)]

#average the OID expression levels by protein and person
averaged_residmat <- residmat_fid %>%
  group_by(FID) %>%
  summarise(across(starts_with("OID"), mean))

averaged_olink <- olink_fid %>%
  group_by(FID) %>%
  summarise(across(starts_with("OID"), mean))

#take out proteins not included in analysis
rows_to_remove <- c("NWD117521", "NWD124541", "NWD159446", "NWD392992", "NWD462959", "NWD523161", "NWD677387", "NWD681170", "NWD857221", "NWD996654")
cols_to_remove <- c("OID30376", "OID31431")

filtered_avg_olink <- averaged_olink[!averaged_olink$FID %in% rows_to_remove, ]
filtered_avg_olink <- filtered_avg_olink[, !colnames(filtered_avg_olink) %in% cols_to_remove]
filtered_avg_residmat <- averaged_residmat[!averaged_residmat$FID %in% rows_to_remove, ]
filtered_avg_residmat <- filtered_avg_residmat[, !colnames(filtered_avg_residmat) %in% cols_to_remove]

#subtract residuals
adjusted_matrix <- filtered_avg_olink[, -1] - filtered_avg_residmat[, -1]
adjusted_matrix <- cbind(filtered_avg_olink[, 1], adjusted_matrix)

#transpose table and assign column and row headers
adjusted_residmat <- t(adjusted_matrix[, -1])
adjusted_residmat <- as.data.table(adjusted_residmat)
colnames(adjusted_residmat) <- adjusted_matrix[,1]
oids<- colnames(adjusted_matrix)[-1]
oids <- as.data.table(oids)
final_residmat <- cbind(oids$oids, adjusted_residmat)
colnames(final_residmat)[1] <- "oids"

#write out
write.table(final_residmat, "~/ancestry_specific_folder/residmat_table", row.names=FALSE, col.names=TRUE, sep="\t", quote=FALSE)

#pre and post lm
#get residuals average, this is the final value compare with random effects 
#plot random effect value vs average residual value (correlation for each protein)
#people v protein pre and post adjustment
```

